---
title: "Penguins"
author: "Connor"
date: "5/31/2023"
output: html_document
---

## Welcome to SPICE 2023

This code chunk is where we load in all of the packages that we will use in this script using library(packagename)

- Packages allow us to interact with our data through the use of functions -> function(data)

- If you get the error: "package "packagename" does not exist":
  - You need to install this package using install.packages("packagename") in the console below

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)# allows us to cut out long file paths (ex. "users/connor/dowloads/etc")
library(tidyverse) # allows us to do lots of fundamental data science tasks (ex. read_csv)
library(ggcorrplot) # allows us to make correlation plots
library(plotly) # allows us to make ggplot objects interactive
library(janitor) # allows us to clean the column names of our dataframe
```






Here is where we read in our data

- We use read_csv from the tidyverse because the file (dataset) we want to read in is a csv (comma separated values)

- We use here to cut out the long file paths and start within our R project
  - Within our project, the data is in the data folder so we direct our computer to the file with "data/penguins_data/dataset_name"

```{r}
penguins_data <- read_csv(here("AFRL_Fall_2023/penguins/penguins_lter.csv"))
```

Clean column names of the dataframe using the clean_names function from the janitor package

- The %>% is called a "pipe". This allows us to make a new dataframe from an existing dataframe and apply new analysis techniques
  - The keyboard shortcut for a pipe is "command + shift + i"

```{r}
penguins_data <- penguins_data %>% 
  clean_names()
```



Exploratory Analysis

Let's begin by exploring some of the columns of our dataset

There is an island column 
- Let's see how many different islands there are in the dataset and how many penguins from the study are on each island
  - We can use the 'table' function to accomplish this
  
```{r}
table(penguins_data$island)
```


Now, let's look at the body mass column by creating a histogram

```{r}
ggplot(data = penguins_data, aes(x = body_mass_g, fill = species)) +
  geom_histogram()
```

```{r}
ggplot(data = penguins_data, aes(x = body_mass_g,
                                 y = delta_15_n_o_oo,
                                 color = species,
                                 shape = island)) +
  geom_point()
```










This looks good, now let's clean this up and look at body mass by species of penguin

```{r}
ggplot(data = penguins_data, aes(x = body_mass_g,
                                 fill = species)) +
  geom_histogram(color = "black") +
  theme_minimal() +
  labs(title = "Penguins, Palmer Station LTER",
       subtitle = "Body Mass Distribution for Adelie, Chinstrap and Gentoo Penguins",
       x = "Body mass (g)",
       y = "Number of Penguins",
       color = "Penguin species")
  
```

Now let's look at the relationship between flipper length and body mass with a scatter plot

```{r}
ggplot(data = penguins_data, aes(x = flipper_length_mm, 
                                 y = body_mass_g
                                 )) +
         geom_point() +
          geom_smooth() +
          theme_minimal() +
  labs(title = "Relationship between Body Mass and Flipper Length",
       x = "Flipper Lenghts (mm)",
       y = "Body Mass")
```

```{r}
ggplot(data = penguins_data, aes(x = culmen_depth_mm,
                                 y = flipper_length_mm,
                                 color = species)) +
  geom_point() +
  geom_smooth()
```

Bar chart of averagebody mass by species

```{r}
average_body_mass_by_species <- penguins_data %>%
  group_by(species) %>% 
  summarize(average_body_mass_g = mean(body_mass_g, na.rm = TRUE))
```

```{r}
ggplot(data = average_body_mass_by_species, aes(y = species, 
                                                x = average_body_mass_g)) +
  geom_col(stat = "identity")
```













```{r}
weights <- 10 + 15 + 28
weights
```



```{r}
mean(weights)
```


It looks like there is a strong positive correlation between body mass and flipper length
- Let's add some variables from the dataset
  - Island
  - Species
  - Sex


```{r}
ggplot(data = penguins_data, aes(x = flipper_length_mm, 
                                 y = body_mass_g,
                                 color = species,
                                 shape = sex)) +
         geom_point() +
  theme_minimal() +
  labs(title = "Penguin size, Palmer Station LTER",
       subtitle = "Flipper length and body mass for Adelie, Chinstrap and Gentoo Penguins",
       x = "Flipper length (mm)",
       y = "Body mass (g)",
       color = "Penguin species",
       shape = "Penguin sex") +
  theme(axis.text.x = element_text(angle = 45)) +
  facet_grid(~island)
```

Now let's explore the relationships between our numeric columns in the dataset with a correlation matrix

Select numeric columns for correlations

```{r}
penguins_data_numeric <- penguins_data %>% 
  select(culmen_length_mm, culmen_depth_mm, flipper_length_mm, body_mass_g)
```

Create a correlation matrix

```{r}
cor_matrix <- cor(penguins_data_numeric[complete.cases(penguins_data_numeric), ], use = "pairwise.complete.obs")
```

Plot the correlation matrix

```{r}
corrplot <- ggcorrplot(cor_matrix, type = "lower", outline.color = "white") +
  theme(axis.text.x = element_text(size = 3),
        axis.text.y = element_text(size = 3))
corrplot
```
Make the correlation plot interactive


```{r}
ggplotly(corrplot)
```
