---
title: "Aggregation and Clustering"
author: "Connor Flynn"
date: "9/7/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(tidyverse)
library(janitor)
library(naniar)
library(plotly)
options(scipen = 99)
```

Read in Data

```{r, message=FALSE, warning=FALSE, include=FALSE}
nano_e_tox <- read_csv(here("data/Beilstein_J_Nanotechnol-06-1788-s002/NanoE-Tox-database.csv"))
```

Select columns up to column 35 (the rest are blank)

```{r}
nano_e_tox <- nano_e_tox[, 1:35]
```

Clean Column Names

```{r}
nano_e_tox <- nano_e_tox %>% 
  clean_names()
```

Missingness 

```{r}
gg_miss_var(nano_e_tox)
```


Explore the toxicity value column

```{r}
class(nano_e_tox$toxicity_value)
```

Change the class of toxicity_value to factor

```{r}
nano_e_tox$toxicity_value <- as.factor(nano_e_tox$toxicity_value)
```


EC = Effect Concentration
LC = Lethal Concentration
NOEC = No Observed Effect Concentration
LOEC = Lowest Observed Effect Concentration
IC = Inhibitory Concentration


Create a function to aggregate with ChatGPT

```{r}
categorize_toxicity_measure <- function(measure) {
  if (grepl("^E|^L|^I", measure)) {
    return("effect")
  } else if (grepl("^N", measure)) {
    return("no_effect")
  } else {
    return(measure)
  }
}
```


```{r}
aggregated_nano_e_tox <- nano_e_tox

# Apply the categorization function to the toxicity_measure column
aggregated_nano_e_tox$toxicity_measure <- sapply(nano_e_tox$toxicity_measure, categorize_toxicity_measure)
```


```{r}
table(aggregated_nano_e_tox$toxicity_measure)
```
```{r}
effect_aggregate <- c("effect", "no_effect")

aggregated_nano_e_tox <- aggregated_nano_e_tox %>% 
  filter(toxicity_measure %in% effect_aggregate)
```


```{r}
table(aggregated_nano_e_tox$type_of_test_organism_selection_bacteria_algae_crustacean_protozoa_fish_earthworm_other)
```



```{r}
table(aggregated_nano_e_tox$test_endpoint)
```



```{r}
ggplot(aggregated_nano_e_tox, aes(x = name, fill = toxicity_measure)) +
  geom_bar() +
  coord_flip() +
  theme_minimal()
```


What nanomaterials are acting the most similarly in this dataset? 
- Can this data answer that question?
