---
title: "Aggregation and Test Organisms Prone to Effect"
author: "Connor Flynn + Maverick Kaopio"
date: "9/7/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(tidyverse)
library(janitor)
library(naniar)
library(plotly)
library(dplyr)
library(ggpubr)
library(caret) # predictive modeling
library(ggplot2)
library(purrr)
library(randomForest)  # For RandomForest classifier
options(scipen = 99)
```

Read in Data

```{r, message=FALSE, warning=FALSE, include=FALSE}
nano_e_tox <- read_csv(here("data/Beilstein_J_Nanotechnol-06-1788-s002/NanoE-Tox-database.csv"))
```

Select columns up to column 35 (the rest are blank)

```{r}
nano_e_tox <- nano_e_tox[, 1:35]
```

Clean Column Names

```{r}
nano_e_tox <- nano_e_tox %>% 
  clean_names()
```

Missingness 

```{r}
gg_miss_var(nano_e_tox)
```


Explore the toxicity value column

```{r}
class(nano_e_tox$toxicity_value)
```

Change the class of toxicity_value to factor

```{r}
nano_e_tox$toxicity_value <- as.factor(nano_e_tox$toxicity_value)
```


EC = Effect Concentration
LC = Lethal Concentration
NOEC = No Observed Effect Concentration
LOEC = Lowest Observed Effect Concentration
IC = Inhibitory Concentration


Create a function to aggregate with ChatGPT

```{r}
categorize_toxicity_measure <- function(measure) {
  if (grepl("^E|^L|^I", measure)) {
    return("effect")
  } else if (grepl("^N", measure)) {
    return("no_effect")
  } else {
    return(measure)
  }
}
```


```{r}
aggregated_nano_e_tox <- nano_e_tox

# Apply the categorization function to the toxicity_measure column
aggregated_nano_e_tox$toxicity_aggregate <- sapply(nano_e_tox$toxicity_measure, categorize_toxicity_measure)
```


```{r}
aggregated_nano_e_tox %>% 
  select(toxicity_aggregate, toxicity_measure)
```


```{r}
table(aggregated_nano_e_tox$toxicity_aggregate)
```


```{r}
effect_aggregate <- c("effect", "no_effect")

aggregated_nano_e_tox <- aggregated_nano_e_tox %>% 
  filter(toxicity_aggregate %in% effect_aggregate)
```


```{r, include=FALSE}
table(aggregated_nano_e_tox$type_of_test_organism_selection_bacteria_algae_crustacean_protozoa_fish_earthworm_other)
```



```{r, include=FALSE}
table(aggregated_nano_e_tox$test_endpoint)
```



```{r}
ggplot(aggregated_nano_e_tox, aes(x = name, fill = toxicity_aggregate)) +
  geom_bar() +
  coord_flip() +
  theme_minimal()
```


What nanomaterials are acting the most similarly in this dataset? 
- Can this data answer that question?





*What is the test endpoint with the highest vulnerability for each nanomaterial?*

Seperate numeric values and units into 2 separate columns
- value
- unit

Raw values only 

```{r}
aggregated_nano_e_tox_raw_values <- subset(aggregated_nano_e_tox, !grepl("\\(|±|-|>|<|≤|%|~", toxicity_value))
```


```{r}
aggregated_nano_e_tox_raw_values <- aggregated_nano_e_tox_raw_values %>% 
  mutate(value = str_extract(toxicity_value, "[0-9.]+([ ][0-9.]+)?"),
         unit = str_extract(toxicity_value, "[^0-9.]+"))
```


```{r}
aggregated_nano_e_tox_raw_values %>% 
  select(toxicity_value, value, unit)
```

```{r}
table(aggregated_nano_e_tox_raw_values$unit)
```

```{r}
aggregated_nano_e_tox_raw_values$unit <- trimws(aggregated_nano_e_tox_raw_values$unit)
```

```{r}
aggregated_nano_e_tox_raw_values_ml <- aggregated_nano_e_tox_raw_values %>% 
  filter(unit == "mg/L")
```



```{r}
result <- aggregated_nano_e_tox_raw_values_ml %>%
  filter(toxicity_aggregate == "effect") %>%
  group_by(name) %>%
  summarize(
    min_value = min(value, na.rm = TRUE),
    test_organism = first(test_organism)  # Add this line to get the test_organism value
  )
result
```


Heat Map

```{r}
aggregated_nano_e_tox_raw_values_ml %>% 
  select(name, test_organism, value, toxicity_aggregate)
```

```{r}
# Assuming 'value' is not numeric, convert it to numeric
aggregated_nano_e_tox_raw_values_ml$value <- as.numeric(aggregated_nano_e_tox_raw_values_ml$value)

```



```{r}
ggplot(data = aggregated_nano_e_tox_raw_values_ml, aes(x = value)) +
  geom_histogram(bins = 300)
```
Mav input***

```{r}
# Arrange the data by toxicity_aggregate and value in ascending order
# Then, add a rank column based on value within each sample_organism grouping
toxicity_category_list_full <- aggregated_nano_e_tox_raw_values_ml %>%
  filter(toxicity_aggregate == "effect") %>% 
  group_by(test_organism) %>% 
  arrange(test_organism, value) %>%
  mutate(rank = rank(value)) %>%
  select(test_organism, name, value, rank, toxicity_aggregate) %>% 
  mutate(rank = factor(rank)) # convert to discrete from continuous

# Display the resulting dataframe
toxicity_category_list <- toxicity_category_list_full %>%
  filter(rank == 1 | rank == 1.5)

toxicity_category_list
```
Same in philosopy to connor; just different method. Gonna toss anything thats > 1.5 ; keeping 1, and points of contention between chemicals @ same dose.
```{r}
plot <- ggplot(data = toxicity_category_list, aes(x = name , y = test_organism, label = value, fill = rank)) +
  geom_point(size = 1, shape = 21) +
  geom_smooth() +
  labs(title = "Deadliness scale (dose)",
       x = "Toxicity",
       y = "Organism") +
  theme(axis.text.y = element_text(size = 4))

plotly_plot <- ggplotly(plot)
plotly_plot
```
^ ranking based on toxicity concentration; now we should try to establish relative toxicity? because we may have some chemicals that were just not tested on an organism? perhaps an opportunity to use machine learning to predict if a chemical might damage another one in a similar family? idk.

```{r}
plot_2 <- ggplot(toxicity_category_list, aes(x = rank , y = test_organism, fill = log(value), label = value)) +
  geom_tile() +
  facet_wrap(~name) +
  theme(legend.position = "none") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 1)) +
  scale_fill_gradient(low = "blue", high = "red")

plotly_plot <- ggplotly(plot_2)
plotly_plot
```
From this, we can blatantly see that some chemicals are very much more dangerous than others based on the data that we have; albeit this is relative to one another. We could make the assumption that because the poison makes the dose, there should be toxicity for all chemicals; difference being just the scale of the dose; something that can be emulated using machine learning perhaps?

Recombine this new data with the old data as a merge; just want the rank.
```{r}
aggregated_nano_e_tox_raw_values_ml_ranked <- left_join(aggregated_nano_e_tox_raw_values_ml, 
                    toxicity_category_list_full %>% select(rank), 
                    by = "test_organism")
```
^this error is true since I'd used grouping for the test_organism & the chemical & the value of the chemical to determine rank. Added back in all the rankings; not just minimum. might paint a better picture down the road perhaps for chemicals that had an effect, but were just not as effective. Perhaps these have different applications.
```{r}
aggregated_nano_e_tox_raw_values_ml_ranked
```

```{r}
# Filter the data
aggregated_nano_e_tox_raw_values_ml_ranked_outliers <- aggregated_nano_e_tox_raw_values_ml_ranked %>%
  filter(value < 10000)

# Calculate the mean of the "value" column by grouping
aggregated_nano_e_tox_raw_values_ml_ranked_outliers_mean <- aggregated_nano_e_tox_raw_values_ml_ranked_outliers %>% 
  group_by(name, test_organism, toxicity_aggregate) %>%
  summarise(mean_value_dose = mean(as.numeric(value), na.rm = TRUE))

# Display the resulting dataframe
aggregated_nano_e_tox_raw_values_ml_ranked_outliers_mean
```

```{r}
# Fill colors
fill_colors <- c('effect' = 'red', 'no_effect' = 'green') # Would be fun to maybe see if there was some sort of beneficial effect/paint blue?

# Create the initial ggplot plot with updated fill colors
plot <- ggplot(data = aggregated_nano_e_tox_raw_values_ml_ranked_outliers_mean, 
               aes(x = mean_value_dose, y = test_organism, label = name, fill = toxicity_aggregate)) +
  aes(shape = 'test_organism') +
  geom_point(size = 1, shape = 21) +
  geom_smooth() +
  stat_cor(method = "pearson", output.type = "text", label.sep = '\n', label.x = 20) +
  labs(title = "Relationship",
       x = "Toxicity",
       y = "Organism") +
  theme_minimal() +
  scale_fill_manual(values = fill_colors) +
  theme(axis.text.y = element_text(size = 5))  # Adjust the size of y-axis text

ggplotly(plot)

```
Add this back to the OG dataframe

```{r}
aggregated_nano_e_tox_raw_values_ml_ranked_summarized <- left_join(
  aggregated_nano_e_tox_raw_values_ml_ranked,
  aggregated_nano_e_tox_raw_values_ml_ranked_outliers_mean,
  by = c("name","test_organism", "toxicity_aggregate")
)
```
Had chatGPT generate some code to gauge relative size of things to add to the dataframe:
```{r}
# unique(aggregated_nano_e_tox_raw_values_ml_ranked_summarized$test_organism)
# Create a mapping from instance names to size categories
size_mapping <- c(
  "Chlorella sp." = "microscopic",
  "Hyalella azteca" = "small",
  "Chironomus dilutus" = "small",
  "rice" = "small",
  "Amphiascus tenuiremis" = "small",
  "Ceriodaphnia dubia" = "small",
  "Escherichia coli" = "microscopic",
  "P.aeruginosa" = "microscopic",
  "B.subtilis" = "microscopic",
  "S.aureus" = "microscopic",
  "Pseudokirchneriella subcapitata" = "microscopic",
  "Danio rerio" = "small",
  "Oncorhynus mykiss" = "small",
  "Ambystoma mexicanum" = "medium",
  "Xenopus laevis" = "medium",
  "Escherichia coli MG1655" = "microscopic",
  "Cupriavididus metallidurans CH34" = "microscopic",
  "Cucurbita pepo" = "medium",
  "Thalassiosira pseudonana" = "microscopic",
  "Tigriopus  japonicus" = "small",
  "Oryzias  melastigma" = "small",
  "Xenopus larvae" = "medium",
  "Dunaliella tertiolecta" = "microscopic",
  "Chlorella vulgaris" = "microscopic",
  "Oreochromis niloticus" = "small",
  "Daphnia similis" = "small",
  "Mytilus galloprovincialis" = "medium",
  "Daphnia magna" = "small",
  "Bacillus stearothermophilus" = "microscopic",
  "Lemna gibba" = "small",
  "Vibrio fischeri" = "microscopic",
  "Thamnocephalus platyurus" = "small",
  "Lolium perenne" = "medium",
  "Caenorhabditis elegans" = "small",
  "E. coli O111" = "microscopic",
  "Escherichia coli K12 DH10B" = "microscopic",
  "Bodo saltans" = "microscopic",
  "Ceriodaphnia affinis" = "small",
  "Lepidium sativum" = "small",
  "Pseudomonas putida BS566::luxCDABE" = "microscopic",
  "Caenorhabditis elegans Bristol strain N2" = "small",
  "Daphnia pulex" = "small",
  "Chironomus riparius" = "small",
  "Nitzschia palea" = "microscopic",
  "Pleurodeles waltl" = "medium",
  "Cucumis sativus" = "medium",
  "Lactuca sativa" = "medium",
  "anaerobic bacteria" = "microscopic",
  "Pseudomonas putida OUS 82" = "microscopic",
  "Pimephales promelas" = "small",
  "Oncorhynchus mykiss" = "small",
  "Drosophila melanogaster" = "small",
  "flax (Linum usitatissimum L., cv. Electra)" = "medium",
  "ryegrass (Lolium perenne L., cv. Tove)" = "medium",
  "two-rowed barley (Hordeum vulgare L., cv. Annabell)" = "medium",
  "tomato (Lycopersicum esculentum)" = "medium",
  "corn (Zea mays)" = "medium",
  "Pseudomonas chlororaphis O6" = "microscopic",
  "Cucurbita pepo subspecies ovifera (yellow squash)" = "medium",
  "Caenorhabditis elegans mtl-2 mutant" = "small",
  "Caenorhabditis elegans sod-3 mutant" = "small",
  "Caenorhabditis elegans daf-12 mutant" = "small",
  "Lumbricus terrestris" = "medium",
  "Arabidopsis thaliana" = "medium",
  "bacterioplankton" = "microscopic",
  "Eisenia fetida" = "small",
  "Acuatic Chironomus riparius" = "small",
  "Desmodesmus subspicatus" = "microscopic",
  "Cyprinus carpio" = "small",
  "Triticum aestivum spp" = "medium",
  "return activated sludge) from Northwest Wastewater Treatment Plant in Mesa, Arizona" = "medium",
  "Eisenia andrei" = "small",
  "Lemna Minor" = "small",
  "Dreissena polymorpha" = "medium",
  "Triticum aestivum" = "medium",
  "Gammarus fossarum" = "small",
  "Anabaena variabilis" = "microscopic",
  "Haliotis diversicolor supertexta" = "medium",
  "Scenedesmus sp" = "microscopic",
  "Chlorella sp" = "microscopic",
  "Lytechinus pictus" = "medium",
  "Chlamydomonas reinhardtii" = "microscopic",
  "Phaeodactylum tricornutum" = "microscopic",
  "Brachionus plicatilis" = "microscopic",
  "Lemna minor" = "small",
  "Montastraea faveolata" = "medium",
  "Caenorhabditis elegans wild-type Maupus, N2 var. Bristol" = "small",
  "Tetrahymena thermophila" = "microscopic",
  "Allogamus ligonifer" = "small",
  "Schoenoplectus tabernaemontani" = "medium",
  "Cucumis sativus (cucumber)" = "medium",
  "Lactuca sativa (lettuce)" = "medium"
)

# Microscopic: Organisms that are typically too small to be seen without a microscope. Examples include bacteria (Escherichia coli), some algae (Chlorella sp.), and certain microorganisms.

# Small: Organisms that are larger than microscopic but still relatively small in size. Examples include small aquatic organisms (Daphnia magna, Pimephales promelas), small insects (Chironomus dilutus), and small plants (Lepidium sativum).

# Medium: Organisms that are of moderate size and can be seen with the naked eye. Examples include medium-sized aquatic organisms (Mytilus galloprovincialis, Ambystoma mexicanum), some plants (Cucumis sativus), and certain small animals (Gammarus fossarum).

# Large: Organisms that are relatively large in size and easily visible without magnification. Examples include large aquatic organisms (Oncorhynchus mykiss), larger animals (Danio rerio, Xenopus laevis), and medium-sized plants (Arabidopsis thaliana).

# Add the SizeCategory column based on the mapping
aggregated_nano_e_tox_raw_values_ml_ranked_summarized_sized <- 
  aggregated_nano_e_tox_raw_values_ml_ranked_summarized %>%
  mutate(organism_relative_size = size_mapping[test_organism])

aggregated_nano_e_tox_raw_values_ml_ranked_summarized_sized
```
Dose relative to size?

```{r}

plot <- ggplot(data = aggregated_nano_e_tox_raw_values_ml_ranked_summarized_sized, aes(x = name, y = log(value), fill = toxicity_aggregate)) +
  geom_point(size = 1, shape = 21) +
  geom_smooth() +
  labs(title = "",
       x = "Chemical Name",
       y = "Log(Value)") +
  facet_grid(. ~ organism_relative_size) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_manual(values = fill_colors)

ggplotly(plot)
```
any toxicity specific to a species.....?
Need to standardize some of the wording:
```{r}
unique(aggregated_nano_e_tox_raw_values_ml_ranked_summarized_sized$type_of_test_organism_selection_bacteria_algae_crustacean_protozoa_fish_earthworm_other)
# Gotta manually replace the weird stuff
name_replacements <- c("Algae" = "algae",
                       "mussels" = "mussel",
                       "insects" = "insect",
                       "amphipod" = "crustacean",
                       "water flea" = "crustacean",
                       "midge" = "insect",
                       "cyanobacteria" = "algae",
                       "meiobenthic copoepod" = "crustacean",
                       "Zebrafish" = "fish"
                       )

aggregated_nano_e_tox_raw_values_ml_ranked_summarized_sized$type_of_test_organism_selection_bacteria_algae_crustacean_protozoa_fish_earthworm_other <- 
  ifelse(aggregated_nano_e_tox_raw_values_ml_ranked_summarized$type_of_test_organism_selection_bacteria_algae_crustacean_protozoa_fish_earthworm_other %in% names(name_replacements),
         name_replacements[aggregated_nano_e_tox_raw_values_ml_ranked_summarized$type_of_test_organism_selection_bacteria_algae_crustacean_protozoa_fish_earthworm_other],
         aggregated_nano_e_tox_raw_values_ml_ranked_summarized$type_of_test_organism_selection_bacteria_algae_crustacean_protozoa_fish_earthworm_other)

# then regroup things based on larger context since we have the exacts already
map_to_broad_category <- function(term) {
  case_when(
    term %in% c("algae", "plant") ~ "Plants",
    term %in% c("crustacean", "fish", "mussel", "snail", "sea urchin") ~ "Aquatic Animals",
    term %in% c("insect", "amphibian", "earthworm") ~ "Terrestrial Animals",
    term %in% c("bacteria", "nematode", "protozoa", "rotifer") ~ "Microorganisms",
    term == "coral" ~ "Coral",
    TRUE ~ "Other"
  )
}

# make broad categories in a new column
aggregated_nano_e_tox_raw_values_ml_ranked_summarized_sized <- aggregated_nano_e_tox_raw_values_ml_ranked_summarized_sized %>%
  mutate(broad_organism_grouping = map_to_broad_category(type_of_test_organism_selection_bacteria_algae_crustacean_protozoa_fish_earthworm_other))
```

```{r}
plot <- ggplot(data = aggregated_nano_e_tox_raw_values_ml_ranked_summarized_sized, aes(x = name, y = log(value), fill = toxicity_aggregate)) +
  geom_point(size = 1, shape = 21) +
  labs(title = "",
       x = "Chemical Name",
       y = "Log(Value)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_manual(values = fill_colors) +
  facet_wrap(~broad_organism_grouping)

ggplotly(plot)
```

```{r}
# Look at unique names
unique_names <- unique(aggregated_nano_e_tox_raw_values_ml_ranked_summarized_sized$name)

unique_names

# Look at only the carbon based materials; create families for the chemicals similar to above for organisms
carbon_based_materials <- c("MWCNT", "SWCNT", "MWCNT-OH", "MWCNT-COOH", "DWCNT", "C60 fullerene", "fullerene")

aggregated_nano_e_tox_raw_values_ml_ranked_summarized_sized_chem_organised <- aggregated_nano_e_tox_raw_values_ml_ranked_summarized_sized %>%
  mutate(carbon_based = ifelse(name %in% carbon_based_materials, "Carbon-Based", "Metal"))
```
```{r}
# Gonna just look at carbon based stuff for now it looks like
carbon_based <- aggregated_nano_e_tox_raw_values_ml_ranked_summarized_sized_chem_organised %>% filter(name %in% carbon_based_materials)

carbon_based
```
Investigate the properties of fullerene as a stabilizer of cell membranes?


Can Fullerine be used as a clothing additive which prevents bacterial buildup? (since its a photosensitized chemical), which would be negligible if consumed.
Can Fullerene heal cystic fibrosis/ lungs damaged by chemicals of war?







I want to generate predictive data based on the data we have that may indicate if a chemical may be poisonous at a different dose:
name, test_organism, effect, value, rank, mean_value_dose

```{r}
# Removing missing values
missing_values <- colSums(is.na(aggregated_nano_e_tox_raw_values_ml_ranked_summarized_sized[, c('name', 'broad_grouping', 'test_organism', 'value','toxicity_aggregate')]))

# Print the counts of missing values
print(missing_values)

# Remove rows with missing values in 'value' column
aggregated_nano_e_tox_raw_values_ml_ranked_summarized_sized <- aggregated_nano_e_tox_raw_values_ml_ranked_summarized_sized[complete.cases(aggregated_nano_e_tox_raw_values_ml_ranked_summarized_sized[, c('name', 'broad_grouping', 'test_organism', 'value','toxicity_aggregate')]), ]
```

```{r}
# Select specific columns and create a new dataframe 'df'
df <- aggregated_nano_e_tox_raw_values_ml_ranked_summarized_sized[, c('name', 'broad_grouping', 'test_organism', 'value', 'toxicity_aggregate')]

# Set a seed for reproducibility
set.seed(42)

# Split the data into training and testing sets
test_proportion <- 0.2
num_rows <- nrow(df)
num_test <- round(num_rows * test_proportion)

# Generate random indices for the testing set
test_indices <- sample(seq_len(num_rows), size = num_test)

# Create the training and testing data frames
train_df <- df[-test_indices, ]
test_df <- df[test_indices, ]

# Step 3: Train a machine learning model
X_train <- train_df[, c('broad_grouping', 'test_organism', 'name', 'toxicity_aggregate', 'value')]
y_train <- train_df$value

# Create the RandomForest model
model <- randomForest(y_train ~ ., data = train_df, ntree = 100, random_state = 42)

# Step 4: Evaluate the model
X_test <- test_df[, c('broad_grouping', 'test_organism', 'name', 'toxicity_aggregate','value')]
y_test <- test_df$value

# Make predictions using the trained model
y_pred <- predict(model, newdata = X_test)

# Calculate accuracy
accuracy <- sum(y_test == y_pred) / length(y_test)
print(paste("Accuracy on the test set:", round(accuracy, 2)))

# Define a list of test organisms to make predictions
target_organisms <- c("bacteria", "algae", "crustacean", "protozoa", "fish", "earthworm", "other")

# Create an empty list to store predictions for each organism
predictions_list <- list()

# Iterate through each organism and make predictions
for (organism in target_organisms) {
  # Filter the data frame for the current organism
  organism_df <- test_df[test_df$test_organism == organism, c('broad_grouping', 'test_organism', 'name', 'toxicity_aggregate', 'value')]
  
  # Make predictions for the current organism
  predictions <- predict(model, newdata = organism_df)
  
  # Store the predictions in the list
  predictions_list[[organism]] <- predictions
}

predictions_list[['bacteria']]
```

7) Density = size of nanomaterial * Dose? Then Graph that in 3d

Overall Plan:
1) Heatmap project of rank to organisms to find deadliest chemical to dose ratio, and also what is least deadliest
2) then see if I can push the list to GPT or find another data list to add to mine that talks about human toxicity to overlap with what Ali'i was thinking. merge w/ ali'i's consolidation of names concept, then see applicability for antibiotics/ anti-microbial sprays? Just keeping to the bacteria group of course.
3) Would be good to maybe intersect the general toxicity trends/ ranked toxicity lists both into a new dataframe that includes the non-toxic, and denotes non-toxic as all being a null value? or maybe 0 since its a scale of 1, and up?
size of particle * concentration and then designate shape based on its shape again, on the 3d graph?
4) curious about if particles may have a positive effect as well
5) create machine learning model to create new data points for future testing; using individual datapoints and the means (or median better?).
6) 2 dendrograms; one for materials, and the others for organisms, then intersect them?




What nanomaterials are acting the most similarly in this dataset? 
- Can this data answer that question?

Should be able to; just need to:
1) Look at scoring, and the range for each category of effect
2) For each chemical: Classify that impact as low/ med/ high? 1/3:1/3:1/3? (simpler alternative)
OR K-means for each? then select the highest group

3) Do this for all columns
4) Alias and select only the high impact columns/values for each chemical?
5) Rate them based on relative impact to one another/scale, to determine highest to lowest impact.

* There appears to be an error in the encoding of numbers in the diamerter_nm column to dates, and some in length_nm as well

Look at when toxicity_measure = effect, establish the scale of effect, then apply back the statistics of what percentage had and didnt as a weight?

^ Better Classify the data down to the shape and size? How granular does it need to be?
Either way, I can just average all the data between what lived and died, and t-test it all perhaps.



What is the test endpoint with the highest vulnerability for each nanomaterial?

Ideas List:

Simplest idea was just using nanoparticles as a good surface cleaner. Think Ali'i brought that one up. Can't remember.

could we filter water using high zeta potential, low toxicity nanoparticles? maybe in combonation with a solar still? figure out what size/kinds of particles dont transfer in steam vapor?

https://link.springer.com/article/10.1007/s10311-023-01626-x
https://www.sciencedirect.com/science/article/pii/S0048969721049135



Maybe nano particles can be paired to help remove one another from a solvent? just used to clean something, then 

How about a better smoke grenade that works against thermal from jets? Look at stuff thats not toxic to humans in large doses, then determine the refractive index of the nano particles? Perhaps that also will not burn? then look at the high zeta materials which would distribute well in a mix, then determine what a mix would need. "Aerosol of burning phosphorus particles is an effective obscurant against thermal imagingsystems. However, this effect is short-lived. After the phosphorus particles fully burn." hopefully the dust wouldnt cause an explosion though. or maybe it would and we have a better explosive lol. could be coated to be made inert. Could focus on organic vs ionic chemicals. non-ionic should be less reactive to explode...?

Bird malaria preventer?

nano particles in paint? deflection of UV?

Is there a microparticle that could help attract micro plastics? like an ionic that mixes with the non-ionic and shifts the zeta potential so that electrical current could pick it all up to be collected and filtered from water. then used magnet to filter the plastic/ particles away? how to seperate?

How about nano-particle embedded clothing that reflects heat/UV without increasing the brightness of the person/ clothing? Options: Insulation to match environment or reflection.

maybe a nanoparticle mix in clothes that would increase the speed of evaporation of water from clothes?

Med-gel? Some sort of mixture of an adhesive, nanoparticles, and a coagulant that effectively when mixed like epoxy causes a very mild chemical, electrical reaction. Maybe it starts as a well spread mixture, then the chemicals react to change the zeta of the nanoparticles to clump? Then those nano particles dissolve into the tissue. (Nano is literally smaller than bloodcells.) Ionic nano particles are mixed into a gel and the gel fills the wound. The nanoparticles react with blood cells and other organic tissue and forms a protein corona. An electric field is created by chemical reaction thats causes a clumping effect by changing the zeta of the nanomaterials in the corona that then effectively seals fills the gel/scaffold area with coagulated blood allowing for efficient healing.

^simpler version: Use nanoparticles that are dissolvable to create protien coronas that facilitate blood clotting while providing a scaffold of sorts to heal in? (gel) want low zeta potential to clump, but perhaps a high zeta somehow when stored to ensure efficient mix in the gel?

what bio-physics or chemical based properties which have numbers or specific qualities that could be categorized would be criteria for such a tool?

How about nanoparticles which have a beneficial effect? Perhaps can locally detox things like lung cells from damage or something by providing substrate directly to the cell? or perhaps helping the cell to rid itself of some toxic chemical? How about to help cystic fibrosis?

Antimicrobials? Maybe a nanoparticle that can be placed into the clothing somehow...?
