---
title: "Aggregation and Test Organisms Prone to Effect"
author: "Connor Flynn"
date: "9/7/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(tidyverse)
library(janitor)
library(naniar)
library(plotly)
options(scipen = 99)
```

Read in Data

```{r, message=FALSE, warning=FALSE, include=FALSE}
nano_e_tox <- read_csv(here("data/Beilstein_J_Nanotechnol-06-1788-s002/NanoE-Tox-database.csv"))
```

Select columns up to column 35 (the rest are blank)

```{r}
nano_e_tox <- nano_e_tox[, 1:35]
```

Clean Column Names

```{r}
nano_e_tox <- nano_e_tox %>% 
  clean_names()
```

Missingness 

```{r}
gg_miss_var(nano_e_tox)
```


Explore the toxicity value column

```{r}
class(nano_e_tox$toxicity_value)
```

Change the class of toxicity_value to factor

```{r}
nano_e_tox$toxicity_value <- as.factor(nano_e_tox$toxicity_value)
```


EC = Effect Concentration
LC = Lethal Concentration
NOEC = No Observed Effect Concentration
LOEC = Lowest Observed Effect Concentration
IC = Inhibitory Concentration


Create a function to aggregate with ChatGPT

```{r}
categorize_toxicity_measure <- function(measure) {
  if (grepl("^E|^L|^I", measure)) {
    return("effect")
  } else if (grepl("^N", measure)) {
    return("no_effect")
  } else {
    return(measure)
  }
}
```


```{r}
aggregated_nano_e_tox <- nano_e_tox

# Apply the categorization function to the toxicity_measure column
aggregated_nano_e_tox$toxicity_aggregate <- sapply(nano_e_tox$toxicity_measure, categorize_toxicity_measure)
```


```{r}
aggregated_nano_e_tox %>% 
  select(toxicity_aggregate, toxicity_measure)
```


```{r}
table(aggregated_nano_e_tox$toxicity_aggregate)
```


```{r}
effect_aggregate <- c("effect", "no_effect")

aggregated_nano_e_tox <- aggregated_nano_e_tox %>% 
  filter(toxicity_aggregate %in% effect_aggregate)
```


```{r, include=FALSE}
table(aggregated_nano_e_tox$type_of_test_organism_selection_bacteria_algae_crustacean_protozoa_fish_earthworm_other)
```



```{r, include=FALSE}
table(aggregated_nano_e_tox$test_endpoint)
```



```{r}
ggplot(aggregated_nano_e_tox, aes(x = name, fill = toxicity_aggregate)) +
  geom_bar() +
  coord_flip() +
  theme_minimal()
```


What nanomaterials are acting the most similarly in this dataset? 
- Can this data answer that question?





*What is the test endpoint with the highest vulnerability for each nanomaterial?*

Seperate numeric values and units into 2 separate columns
- value
- unit

Raw values only 

```{r}
aggregated_nano_e_tox_raw_values <- subset(aggregated_nano_e_tox, !grepl("\\(|±|-|>|<|≤|%|~", toxicity_value))
```


```{r}
aggregated_nano_e_tox_raw_values <- aggregated_nano_e_tox_raw_values %>% 
  mutate(value = str_extract(toxicity_value, "[0-9.]+([ ][0-9.]+)?"),
         unit = str_extract(toxicity_value, "[^0-9.]+"))
```


```{r}
aggregated_nano_e_tox_raw_values %>% 
  select(toxicity_value, value, unit)
```

```{r}
table(aggregated_nano_e_tox_raw_values$unit)
```

```{r}
aggregated_nano_e_tox_raw_values$unit <- trimws(aggregated_nano_e_tox_raw_values$unit)
```

```{r}
aggregated_nano_e_tox_raw_values_ml <- aggregated_nano_e_tox_raw_values %>% 
  filter(unit == "mg/L")
```



```{r}
result <- aggregated_nano_e_tox_raw_values_ml %>%
  filter(toxicity_aggregate == "effect") %>%
  group_by(name) %>%
  summarize(
    min_value = min(value, na.rm = TRUE),
    test_organism = first(test_organism)  # Add this line to get the test_organism value
  )
result
```


Heat Map

```{r}
aggregated_nano_e_tox_raw_values_ml %>% 
  select(name, test_organism, value, toxicity_aggregate)
```

```{r}

# Assuming 'value' is not numeric, convert it to numeric
aggregated_nano_e_tox_raw_values_ml$value <- as.numeric(aggregated_nano_e_tox_raw_values_ml$value)

# Your ggplot code
ggplot(aggregated_nano_e_tox_raw_values_ml, aes(x = name, y = test_organism, fill = value)) +
  geom_tile() +
  facet_wrap(~toxicity_aggregate) +
  theme(legend.position = "none") +
  theme_minimal() +
  
  # Specify a continuous color scale
  scale_fill_gradient(low = "blue", high = "red")  # You can customize the colors as needed

```
```{r}
ggplot(data = aggregated_nano_e_tox_raw_values_ml, aes(x = value)) +
  geom_histogram(bins = 300)
```



Alii Idea 

```{r}
test_orgs <- as.data.frame(table(aggregated_nano_e_tox_raw_values_ml$test_organism))
```

```{r}
type_test_orgs <- as.data.frame(table(aggregated_nano_e_tox_raw_values_ml$type_of_test_organism_selection_bacteria_algae_crustacean_protozoa_fish_earthworm_other))
```


```{r}
bacteria <- aggregated_nano_e_tox_raw_values_ml %>% 
  filter(type_of_test_organism_selection_bacteria_algae_crustacean_protozoa_fish_earthworm_other == "bacteria")
```

```{r}
bacteria_test_orgs <- as.data.frame(table(bacteria$test_organism))
```

