---
title: "NanoE-Tox Exploratory Analysis"
author: "Connor Flynn"
date: "7/10/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(tidyverse)
library(janitor)
library(naniar)
options(scipen = 99)
```

Read in Data

```{r}
nano_e_tox <- read_csv(here("data/Beilstein_J_Nanotechnol-06-1788-s002/NanoE-Tox-database.csv"))
```

Select columns up to column 35 (the rest are blank)

```{r}
nano_e_tox <- nano_e_tox[, 1:35]
```

Clean Column Names

```{r}
nano_e_tox <- nano_e_tox %>% 
  clean_names()
```

Missingness 

```{r}
gg_miss_var(nano_e_tox)
```


Explore the toxicity value column

```{r}
class(nano_e_tox$toxicity_value)
```

Change the class of toxicity_value to factor

```{r}
nano_e_tox$toxicity_value <- as.factor(nano_e_tox$toxicity_value)
```

Find the unique values in the toxicity_value column

```{r}
unique_toxicity_values <- unique(nano_e_tox$toxicity_value)
```

Store the unique values as levels in a dataframe

```{r}
levels_unique_toxicity_values <- as.data.frame(levels(unique_toxicity_values))
```




**Challenge** Normalizing

Exploration pre-normalizing 

1. how many ranges (-)

```{r}
toxicity_ranges <- subset(nano_e_tox, grepl("-", toxicity_value))
```




2. how many greater than less than (> <)

```{r}
toxicity_greater_less <- subset(nano_e_tox, grepl("[><]", toxicity_value))
```

3. how many ±

```{r}
toxicity_plus_minus <- subset(nano_e_tox, grepl("[±]", toxicity_value))
```

4. how many confidence intervals (%)

```{r}
toxicity_confidence <- subset(nano_e_tox, grepl("[%]", toxicity_value))
```

Raw values only 

```{r}
nano_raw_toxicity_values <- subset(nano_e_tox, !grepl("\\(|±|-|>|<|≤|%", toxicity_value))
```



Seperate numeric values and units into 2 separate columns
- value
- unit


```{r}
nano_raw_toxicity_values <- nano_raw_toxicity_values %>% 
  mutate(value = str_extract(toxicity_value, "[0-9.]+([ ][0-9.]+)?"),
         unit = str_extract(toxicity_value, "[^0-9.]+"))
```


```{r}
nano_raw_toxicity_values %>% 
  select(toxicity_value, value, unit)
```

Find unique values in the unit column

```{r}
unique(nano_raw_toxicity_values$unit)
```

```{r}
nano_raw_toxicity_values$value <- as.numeric(nano_raw_toxicity_values$value)
```



Create function

```{r}
convert_to_mg_L <- function(df) {
  unit_conversion <- function(value, unit) {
    if (is.na(unit) || is.na(value)) {
      return(NA)
    } else {
      unit <- trimws(tolower(unit))
      
      if (unit == "mg/l" || unit == "mg zn/kg d" || unit == "mg/kg" || unit == "mg/kg dry food" ||
          unit == "mg zn/kg" ||  unit == "mg zn/ml" || unit == "mg/ml") {
        return(value)
      } else if (unit == "g/l" || unit == "g/kg" || unit == "mg/g" || unit == "mg/g dry matter") {
        return(value * 1000)
      } else if (unit == "µg/ml" || unit == "µg/g of food" || unit == "µg/l" || unit == "µg dissolved zn/l" ||
                 unit == "µg ag/l" || unit == "µg dissolved ag/l" || unit == "µg" || unit == "µg/g" ||
                 unit == "µg zn/g" || unit == "µg zn/mg" || unit == "µg zn/ml" || unit == "µg cu/g") {
        return(value * 0.001)
      } else if (unit == "ppm" || unit == "mg ag/l" || unit == "nm" || unit == "nmol/g fresh weight") {
        return(value)
      } else if (unit == "mmol/g" || unit == "µm" || unit == "m" || unit == "molar") {
        return(NA)
      } else {
        return(NA)
      }
    }
  }
  
  df$standardized_values <- unit_conversion(df$value, df$unit)
  return(df)
}



```

```{r}
nano_raw_toxicity_values <- convert_to_mg_L(nano_raw_toxicity_values)
```


```{r}
nano_raw_toxicity_values %>% 
  select(toxicity_value, value, unit, standardized_values)
```
Results Test

mg/L to mg/L doesn't work
- 1 mg/mL = 1 mg/mL

mg/mL to mg/L works
- 1 mg/mL = 1,000 mg/L

g/l to mg/L works 
- 1 g/l = 1,000 mg/L

µg/mL to mg/L doesn't work 
- 1 µg/mL = 1 mg/L

```{r}

```








```{r}
nano_e_tox <- nano_e_tox %>%
  mutate(value = str_extract(toxicity_value, "\\(?([0-9.]+)\\s*-?\\s*[0-9.]*\\)?|([0-9.]+)\\s*±\\s*[0-9.]+"),
         unit = str_extract(toxicity_value, "(?<=\\s)[^0-9.()\\s±]+(?=\\s|$)"))
```

Check results

```{r}
nano_e_tox %>% 
  select(toxicity_value, value, unit)
```


Define a function to convert units to mg/L
```{r}
convert_to_mg_L <- function(value, unit) {
  if (unit == "mg/L") {
    return(value)
  } else if (unit == "mg/kg") {
    return(value)
  } else if (unit == "nM") {
    # Conversion factor: 1 nM = 1e-6 mg/L
    return(value * 1e-6)
  } else if (unit == "ppm") {
    # Conversion factor: 1 ppm = 1 mg/L
    return(value)
  } else if (unit == "ppt (ng/L?)") {
    # Conversion factor: 1 ppt = 1 ng/L = 1e-6 mg/L
    return(value * 1e-6)
  } else if (unit == "µg dissolved Zn/L" ||
             unit == "µg/L" ||
             unit == "µg/g" ||
             unit == "µg/mL" ||
             unit == "µM" ||
             unit == "µg" ||
             unit == "µg/g of food") {
    # Conversion factor: 1 µg = 1e-3 mg
    return(value * 1e-3)
  } else if (unit == "M ZnO") {
    # Conversion factor: 1 M = 1e6 mg/L
    return(value * 1e6)
  } else if (unit == "mg Cu/L") {
    return(value)
  } else if (unit == "mg/kg of dry feed" ||
             unit == "mg/kg dry food" ||
             unit == "mg/kg" ||
             unit == "mg" ||
             unit == "mg/g") {
    # Conversion factor: 1 mg = 1 mg/L
    return(value)
  } else {
    return(NA)
  }
}
```


Apply the conversion function to normalize the values

```{r}
nano_e_tox <- nano_e_tox %>%
  mutate(normalized_toxicity_value = convert_to_mg_L(value, str_extract(toxicity_value, "[^0-9.]+")))
```







