---
title: "NanoE-Tox Exploratory Analysis"
author: "Connor Flynn"
date: "7/10/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(tidyverse)
library(janitor)
library(naniar)
```

Read in Data

```{r}
nano_e_tox <- read_csv(here("data/Beilstein_J_Nanotechnol-06-1788-s002/NanoE-Tox-database.csv"))
```

Select columns up to column 35 (the rest are blank)

```{r}
nano_e_tox <- nano_e_tox[, 1:35]
```

Clean Column Names

```{r}
nano_e_tox <- nano_e_tox %>% 
  clean_names()
```

Missingness 

```{r}
gg_miss_var(nano_e_tox)
```


Explore the toxicity value column

```{r}
class(nano_e_tox$toxicity_value)
```

Change the class of toxicity_value to factor

```{r}
nano_e_tox$toxicity_value <- as.factor(nano_e_tox$toxicity_value)
```

Find the unique values in the toxicity_value column

```{r}
unique_toxicity_values <- unique(nano_e_tox$toxicity_value)
```

Store the unique values as levels in a dataframe

```{r}
levels_unique_toxicity_values <- as.data.frame(levels(unique_toxicity_values))
```




**Challenge** Normalizing

Exploration pre-normalizing 

1. how many ranges (-)

```{r}
toxicity_ranges <- subset(nano_e_tox, grepl("-", toxicity_value))
```

```{r}
toxicity_ranges <- toxicity_ranges[!grepl("\\([^\\)]+\\)", toxicity_ranges$toxicity_value), ]
```


2. how many greater than less than (> <)

```{r}
toxicity_greater_less <- subset(nano_e_tox, grepl("[><]", toxicity_value))
```

3. how many ±

```{r}
toxicity_plus_minus <- subset(nano_e_tox, grepl("[±]", toxicity_value))
```

4. how many confidence intervals (%)

```{r}
toxicity_confidence <- subset(nano_e_tox, grepl("[%]", toxicity_value))
```

Raw values only 

```{r}
nano_raw_toxicity_values <- subset(nano_e_tox, !grepl("\\(|±|-|>|<|≤|%", toxicity_value))
```



Seperate numeric values and units into 2 separate columns
- value
- unit


```{r}
nano_raw_toxicity_values <- nano_raw_toxicity_values %>% 
  mutate(value = str_extract(toxicity_value, "[0-9.]+([ ][0-9.]+)?"),
         unit = str_extract(toxicity_value, "[^0-9.]+"))
```


```{r}
nano_raw_toxicity_values %>% 
  select(toxicity_value, value, unit)
```

Find unique values in the unit column

```{r}
unique(nano_raw_toxicity_values$unit)
```




Create function

```{r}
standardize_to_mg_per_L <- function(value, unit) {
  # Convert value to numeric
  value <- as.numeric(value)
  
  # Standardize based on unit
  if (is.na(unit)) {
    # If unit is NA, return value as it is
    standardized <- value
  } else {
    unit <- trimws(unit)  # Remove leading/trailing whitespace
    
    if (grepl("µg/mL", unit) || grepl("ppm", unit)) {
      # Convert µg/mL or ppm to mg/L
      standardized <- value * 1000
    } else if (grepl("mg/kg", unit) || grepl("µg/g", unit) || grepl("mg Ag/kg", unit) || grepl("mg Zn/kg", unit) || grepl("mg/g", unit) || grepl("mg Cu/g", unit)) {
      # Convert mg/kg, µg/g, mg Ag/kg, mg Zn/kg, mg/g, mg Cu/g to mg/L
      standardized <- value
    } else if (grepl("g/l", unit) || grepl("g/L", unit)) {
      # Convert g/l or g/L to mg/L
      standardized <- value * 1000
    } else if (grepl("µg/L", unit) || grepl("mg/ml", unit)) {
      # Convert µg/L or mg/ml to mg/L
      standardized <- value
    } else if (grepl("µg dissolved", unit) || grepl("µg Ag/L", unit) || grepl("µg Cu/g", unit) || grepl("nmol/g fresh weight", unit) || grepl("nM", unit)) {
      # Convert µg dissolved, µg Ag/L, µg Cu/g, nmol/g fresh weight, nM to mg/L
      standardized <- value / 1000
    } else if (grepl("mmol/g", unit) || grepl("mM", unit)) {
      # Convert mmol/g or mM to mg/L
      standardized <- value * 1000
    } else if (grepl("mg Zn/kg", unit)) {
      # Convert mg Zn/kg to mg/L
      standardized <- value
    } else {
      # If unit is not recognized, return value as it is
      standardized <- value
    }
  }
  
  return(standardized)
}

# Apply the function to create the standardized_values column
nano_raw_toxicity_values <- nano_raw_toxicity_values %>%
  mutate(standardized_values = format(standardize_to_mg_per_L(value, unit), scientific = FALSE))
```


```{r}
nano_raw_toxicity_values %>% 
  select(toxicity_value, value, unit, standardized_values)
```
Results Test

mg/mL to mg/L works
- 1 mg/mL = 1,000 mg/L

g/l to mg/L works 
- 1 g/l = 1,000 mg/L

µg/mL to mg/L doesn't work 
- 1 µg/mL = 1 mg/L









```{r}
nano_e_tox <- nano_e_tox %>%
  mutate(value = str_extract(toxicity_value, "\\(?([0-9.]+)\\s*-?\\s*[0-9.]*\\)?|([0-9.]+)\\s*±\\s*[0-9.]+"),
         unit = str_extract(toxicity_value, "(?<=\\s)[^0-9.()\\s±]+(?=\\s|$)"))
```

Check results

```{r}
nano_e_tox %>% 
  select(toxicity_value, value, unit)
```


Define a function to convert units to mg/L
```{r}
convert_to_mg_L <- function(value, unit) {
  if (unit == "mg/L") {
    return(value)
  } else if (unit == "mg/kg") {
    return(value)
  } else if (unit == "nM") {
    # Conversion factor: 1 nM = 1e-6 mg/L
    return(value * 1e-6)
  } else if (unit == "ppm") {
    # Conversion factor: 1 ppm = 1 mg/L
    return(value)
  } else if (unit == "ppt (ng/L?)") {
    # Conversion factor: 1 ppt = 1 ng/L = 1e-6 mg/L
    return(value * 1e-6)
  } else if (unit == "µg dissolved Zn/L" ||
             unit == "µg/L" ||
             unit == "µg/g" ||
             unit == "µg/mL" ||
             unit == "µM" ||
             unit == "µg" ||
             unit == "µg/g of food") {
    # Conversion factor: 1 µg = 1e-3 mg
    return(value * 1e-3)
  } else if (unit == "M ZnO") {
    # Conversion factor: 1 M = 1e6 mg/L
    return(value * 1e6)
  } else if (unit == "mg Cu/L") {
    return(value)
  } else if (unit == "mg/kg of dry feed" ||
             unit == "mg/kg dry food" ||
             unit == "mg/kg" ||
             unit == "mg" ||
             unit == "mg/g") {
    # Conversion factor: 1 mg = 1 mg/L
    return(value)
  } else {
    return(NA)
  }
}
```


Apply the conversion function to normalize the values

```{r}
nano_e_tox <- nano_e_tox %>%
  mutate(normalized_toxicity_value = convert_to_mg_L(value, str_extract(toxicity_value, "[^0-9.]+")))
```







