---
title: "NanoE-Tox Exploratory Analysis"
author: "Connor Flynn"
date: "7/10/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(tidyverse)
library(janitor)
library(naniar)
```

Read in Data

```{r}
nano_e_tox <- read_csv(here("data/Beilstein_J_Nanotechnol-06-1788-s002/NanoE-Tox-database.csv"))
```

Select columns up to column 35 (the rest are blank)

```{r}
nano_e_tox <- nano_e_tox[, 1:35]
```

Clean Column Names

```{r}
nano_e_tox <- nano_e_tox %>% 
  clean_names()
```

Missingness 

```{r}
gg_miss_var(nano_e_tox)
```


Explore the toxicity value column

```{r}
class(nano_e_tox$toxicity_value)
```

Change the class of toxicity_value to factor

```{r}
nano_e_tox$toxicity_value <- as.factor(nano_e_tox$toxicity_value)
```

Find the unique values in the toxicity_value column

```{r}
unique_toxicity_values <- unique(nano_e_tox$toxicity_value)
```

Store the unique values as levels in a dataframe

```{r}
levels_unique_toxicity_values <- as.data.frame(levels(unique_toxicity_values))
```


**Challenge** Normalizing

Seperate numeric values and units into 2 separate columns
- value
- unit

```{r}
nano_e_tox <- nano_e_tox %>%
   mutate(value = str_extract(toxicity_value, "[0-9]+(?:\\.[0-9]+)?"),
         unit = str_extract(toxicity_value, "(?<= )[\\D]+"))
```

Check results

```{r}
nano_e_tox %>% 
  select(toxicity_value, value, unit)
```


Define a function to convert units to mg/L
```{r}
convert_to_mg_L <- function(value, unit) {
  if (unit == "mg/L") {
    return(value)
  } else if (unit == "mg/kg") {
    return(value)
  } else if (unit == "nM") {
    # Conversion factor: 1 nM = 1e-6 mg/L
    return(value * 1e-6)
  } else if (unit == "ppm") {
    # Conversion factor: 1 ppm = 1 mg/L
    return(value)
  } else if (unit == "ppt (ng/L?)") {
    # Conversion factor: 1 ppt = 1 ng/L = 1e-6 mg/L
    return(value * 1e-6)
  } else if (unit == "µg dissolved Zn/L" ||
             unit == "µg/L" ||
             unit == "µg/g" ||
             unit == "µg/mL" ||
             unit == "µM" ||
             unit == "µg" ||
             unit == "µg/g of food") {
    # Conversion factor: 1 µg = 1e-3 mg
    return(value * 1e-3)
  } else if (unit == "M ZnO") {
    # Conversion factor: 1 M = 1e6 mg/L
    return(value * 1e6)
  } else if (unit == "mg Cu/L") {
    return(value)
  } else if (unit == "mg/kg of dry feed" ||
             unit == "mg/kg dry food" ||
             unit == "mg/kg" ||
             unit == "mg" ||
             unit == "mg/g") {
    # Conversion factor: 1 mg = 1 mg/L
    return(value)
  } else {
    return(NA)
  }
}
```


Apply the conversion function to normalize the values

```{r}
nano_e_tox <- nano_e_tox %>%
  mutate(normalized_toxicity_value = convert_to_mg_L(value, str_extract(toxicity_value, "[^0-9.]+")))
```







