---
title: "NanoE-Tox Exploratory Analysis"
author: "Connor Flynn"
date: "7/10/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(tidyverse)
library(janitor)
library(naniar)
options(scipen = 99)
```

Read in Data

```{r, message=FALSE, warning=FALSE, include=FALSE}
nano_e_tox <- read_csv(here("data/Beilstein_J_Nanotechnol-06-1788-s002/NanoE-Tox-database.csv"))
```

Select columns up to column 35 (the rest are blank)

```{r}
nano_e_tox <- nano_e_tox[, 1:35]
```

Clean Column Names

```{r}
nano_e_tox <- nano_e_tox %>% 
  clean_names()
```

Missingness 

```{r}
gg_miss_var(nano_e_tox)
```


Explore the toxicity value column

```{r}
class(nano_e_tox$toxicity_value)
```

Change the class of toxicity_value to factor

```{r}
nano_e_tox$toxicity_value <- as.factor(nano_e_tox$toxicity_value)
```

Find the unique values in the toxicity_value column

```{r}
unique_toxicity_values <- unique(nano_e_tox$toxicity_value)
```

Store the unique values as levels in a dataframe

```{r}
levels_unique_toxicity_values <- as.data.frame(levels(unique_toxicity_values))
```




**Challenge** Normalizing

Exploration pre-normalizing 

1. how many ranges (-)

```{r}
toxicity_ranges <- subset(nano_e_tox, grepl("-", toxicity_value))
```




2. how many greater than less than (> <)

```{r}
toxicity_greater_less <- subset(nano_e_tox, grepl("[><]", toxicity_value))
```

3. how many ±

```{r}
toxicity_plus_minus <- subset(nano_e_tox, grepl("[±]", toxicity_value))
```

4. how many confidence intervals (%)

```{r}
toxicity_confidence <- subset(nano_e_tox, grepl("[%]", toxicity_value))
```

Raw values only 

```{r}
nano_raw_toxicity_values <- subset(nano_e_tox, !grepl("\\(|±|-|>|<|≤|%|~", toxicity_value))
```



Seperate numeric values and units into 2 separate columns
- value
- unit


```{r}
nano_raw_toxicity_values <- nano_raw_toxicity_values %>% 
  mutate(value = str_extract(toxicity_value, "[0-9.]+([ ][0-9.]+)?"),
         unit = str_extract(toxicity_value, "[^0-9.]+"))
```


```{r}
nano_raw_toxicity_values %>% 
  select(toxicity_value, value, unit)
```

Find unique values in the unit column

```{r}
unique(nano_raw_toxicity_values$unit)
```

```{r}
nano_raw_toxicity_values$value <- as.numeric(nano_raw_toxicity_values$value)
```


```{r}
as.data.frame(table(nano_raw_toxicity_values$unit))
```




Create a function to standardize 

```{r}
standardize_to_mg_per_l <- function(df) {
  # Define a conversion factor dictionary for different units
  conversion_factors <- c(
    "mg/L" = 1,
    "mg/mL" = 1000,
    "g/l" = 1000,
    "µg/mL" = 1,
    "mg/kg" = 1,
    "µg/g of food" = 1,
    "g/kg" = 1000,
    "mg/kg dry food" = 1,
    "ppm" = 1,
    "mg Zn/kg  d" = 1,
    "µg/L" = 0.001,
    "mmol/g" = 1,
    "mg Zn/kg" = 1,
    "µg dissolved Zn/L" = 1,
    "nM" = 1,
    "µM" = 1000,
    "µg Ag/L" = 1,
    "µg dissolved Ag/L" = 1,
    "mg/kg of dry feed" = 1,
    "mg Ag/L" = 1,
    "µg/g" = 1,
    "mg Ag/kg" = 1,
    "µg" = 1,
    "mM" = 1000,
    "g/L" = 1000,
    "nmol/g fresh weight" = 1,
    "µg Cu/g" = 1
  )
  
  # Create a new column 'standardized_values' to store the standardized values
  df$standardized_values <- NA
  
  # Loop through the dataframe rows and standardize the values
  for (i in seq_len(nrow(df))) {
    unit <- trimws(df$unit[i])  # Trim leading and trailing spaces
    value <- df$value[i]
    
    # Check if the unit exists in the conversion_factors dictionary
    if (!is.na(unit) && unit %in% names(conversion_factors)) {
      df$standardized_values[i] <- value * conversion_factors[unit]
    } else {
      df$standardized_values[i] <- NA
    }
  }
  
  # Return the dataframe with the standardized values
  return(df)
}
```


```{r}
nano_raw_toxicity_values_standardized <- standardize_to_mg_per_l(nano_raw_toxicity_values)
```


```{r}
nano_raw_toxicity_values_standardized %>% 
  select(toxicity_value, value, unit, standardized_values) 

```

```{r}
as.data.frame(table(nano_raw_toxicity_values_standardized$unit))
```














