---
title: "Overview Analysis"
author: "Armando Luna, Brittany Johnson, Connor Flynn"
date: "12/8/2022"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(readr)
library(janitor)
library(tidyr)
library(tidyverse)
library(plotly)
library(RColorBrewer)
library(dplyr)
library(forcats)
library(hrbrthemes)
library(viridis)
library(paletteer)

options(scipen = 99)

```

The purpose of this R Markdown is to perform a preliminary, overview analysis of the dataset from "Application of Bayesian networks in determining nanoparticle-induced cellular outcomes using transcriptomics". 


Import Data


```{r}
nmdata <- read_csv(here("data/Literature_transcriptomics_review_and_data_of_Nanoparticle_Induced_Cellular_Outcomes.csv"))
```

Clean Column Names

```{r}
nmdata <- nmdata %>% 
  clean_names()
```

Replace Column Names with Abbreviations

```{r}

names(nmdata) <- gsub(x = names(nmdata), 
                                        pattern = "unfolded_protein_responses_upr_and_endoplasmatic_reticulum_er_stress", 
                                        replacement = "UPR_ER_Stress")

names(nmdata) <- gsub(x = names(nmdata), 
                                        pattern = "angiogenenesis_responses", 
                                        replacement = "Angiogenesis")

names(nmdata) <- gsub(x = names(nmdata), 
                                        pattern = "metal_ion_responses", 
                                        replacement = "Metal_Ion")

names(nmdata) <- gsub(x = names(nmdata), 
                                        pattern = "inflammation_and_immune_responses", 
                                        replacement = "Inflammation_Immune")

names(nmdata) <- gsub(x = names(nmdata), 
                                        pattern = "cell_adhesion_and_extracellular_matrix_responses", 
                                        replacement = "Cell_Adhesion_ECM")

names(nmdata) <- gsub(x = names(nmdata), 
                                        pattern = "dna_damage_and_repair_responses", 
                                        replacement = "DNA_Damage_Repair")

names(nmdata) <- gsub(x = names(nmdata), 
                                        pattern = "cell_death_and_apoptosis_responses", 
                                        replacement = "Cell_Death_Apoptosis")

names(nmdata) <- gsub(x = names(nmdata), 
                                        pattern = "cell_cycle_and_proliferation_responses", 
                                        replacement = "Cell_Cycle_and_Proliferation")

names(nmdata) <- gsub(x = names(nmdata), 
                                        pattern = "cytoskeleton_organization_responses", 
                                        replacement = "Cytoskeleton_Organization")


  
```

Replace Nanoparticles with Abbreviations

```{r}
nmdata[nmdata == "Ag_NPs"] <- "Ag"
nmdata[nmdata == "Au_NPs"] <- "Au"

nmdata[nmdata == "CuO_NPs"] <- "CuO"
nmdata[nmdata == "Fe2O3_NPs"] <- "Fe203"
nmdata[nmdata == "Polymeric_NPs"] <- "Polymeric"
nmdata[nmdata == "Polystyrene_NPs"] <- "Polystyrene"
nmdata[nmdata == "TiO2_NPs"] <- "TiO2"
nmdata[nmdata == "ZnO_NPs"] <- "ZnO"
```





Select Nanoparticle and Bioassay Columns

```{r}
nmdata_longer <- nmdata %>% 
  select(nanoparticles, Cell_Cycle_and_Proliferation, Cell_Death_Apoptosis, DNA_Damage_Repair, Cell_Adhesion_ECM, Inflammation_Immune, UPR_ER_Stress, Metal_Ion, Angiogenesis, Cytoskeleton_Organization)
```


Pivot Data Longer to have 1 response column

```{r}
nmdata_longer <- nmdata_longer %>% 
  tidyr::pivot_longer(cols = !nanoparticles, names_to = "responses", 
               values_to = "effects")
```

Remove undercores from resposes and effects for visualization cleanliness

```{r}
nmdata_longer  <- nmdata_longer %>%
  mutate(responses = gsub("_", " ", responses),
         effects = gsub("_", " ", effects))
```


Group By Responses and Effects and Count 

```{r}
nmdata_longer_responses <- nmdata_longer %>% 
  group_by(responses, effects) %>% 
  count()
```




Plot Responses and Effects Count


```{r}
ggplot(data = nmdata_longer_responses, aes(fill = effects, y=n, x=responses)) + 
    geom_bar(position="dodge", stat="identity") + 
  coord_flip() + 
  theme_minimal() +
  scale_fill_brewer(palette = "Dark2", direction = -1)
```


Group by Nanoparticles and Effects and Count 

```{r}
nmdata_longer_nanoparticles <- nmdata_longer %>% 
  group_by(nanoparticles, effects) %>% 
  count()
```


Plot Nanoparticles and Effects Count

```{r}
ggplot(data = nmdata_longer_nanoparticles, aes(fill = effects, y=n, x=nanoparticles)) + 
    geom_bar(position="dodge", stat="identity") + 
  coord_flip() + 
  theme_minimal() +
    scale_fill_brewer(palette = "Dark2", direction = -1)
```

Group by Nanoparticles, Responses, and Effects and Count

```{r}
nmdata_longer_nanoparticles_effects <- nmdata_longer %>% 
  group_by(nanoparticles, responses, effects) %>% 
  count()
```


Plot Nanoparticles, Responses, and Effects Count

```{r}
ggplot(data = nmdata_longer_nanoparticles_effects,                
       aes(x = n,
           y = responses,
           fill = nanoparticles)) + 
  geom_bar(stat = "identity",
           position = "stack") +
  facet_grid(~ effects) + 
  #coord_flip() + 
  theme_minimal() +
  scale_fill_brewer(palette="Set3") 

```

```{r}
nmdata_longer_all <- nmdata %>% 
  tidyr::pivot_longer(cols = c(Cell_Cycle_and_Proliferation:Cytoskeleton_Organization), names_to = "responses", 
               values_to = "effects")
```

```{r}
nmdata_longer_all  <- nmdata_longer_all %>%
  mutate(responses = gsub("_", " ", responses),
         effects = gsub("_", " ", effects))
```

```{r}
nmdata_longer_cell_type_effects <- nmdata_longer_all %>% 
  group_by(cell_type, responses, effects) %>% 
  count()
```

Plot by Cell Type

```{r}
ggplot(data = nmdata_longer_cell_type_effects,                         # Draw barplot with grouping & stacking
       aes(x = n,
           y = responses,
           fill = cell_type)) + 
  geom_bar(stat = "identity",
           position = "stack") +
  facet_grid(~ effects) + 
  #coord_flip() + 
  theme_minimal() +
  scale_fill_brewer(palette="Set3") 

```

```{r}
nmdata_longer_tissue_effects <- nmdata_longer_all %>% 
  group_by(tissue, responses, effects) %>% 
  count()
```


Plot by Tissue

```{r}
ggplot(data = nmdata_longer_tissue_effects,                         # Draw barplot with grouping & stacking
       aes(x = n,
           y = responses,
           fill = tissue)) + 
  geom_bar(stat = "identity",
           position = "stack") +
  facet_grid(~ effects) + 
  #coord_flip() + 
  theme_minimal() +
  scale_fill_paletteer_d("colorBlindness::Blue2DarkOrange12Steps")

```
```{r}
ggplot(data = nmdata_longer_tissue_effects,                         # Draw barplot with grouping & stacking
       aes(x = n,
           y = tissue,
           fill = effects)) + 
  geom_bar(stat = "identity",
           position = "stack") +
  facet_wrap(~ responses) + 
  #coord_flip() + 
  theme_minimal() +
    scale_fill_brewer(palette = "Dark2", direction = -1) 

```

Plot by Surface Coatings


```{r}
nmdata_longer_surface_coatings_effects <- nmdata_longer_all %>% 
  group_by(surface_coatings, responses, effects) %>% 
  count()
```


```{r}
ggplot(data = nmdata_longer_surface_coatings_effects,                         # Draw barplot with grouping & stacking
       aes(x = n,
           y = surface_coatings,
           fill = effects)) + 
  geom_bar(stat = "identity",
           position = "stack") +
  facet_grid(~ responses) + 
  #coord_flip() + 
  theme_minimal() +
  scale_fill_brewer(palette = "Dark2", direction = -1) 

```



Violin Plot

Select Columns of Interest

```{r}
nmdata_longer_dose_duration <- nmdata %>% 
  select(nanoparticles, UPR_ER_Stress, Angiogenesis, Metal_Ion, Inflammation_Immune, Cell_Adhesion_ECM, DNA_Damage_Repair, Cell_Death_Apoptosis, Cell_Cycle_and_Proliferation, Cytoskeleton_Organization, exposure_dose_mg_ml, exposure_duration_h)
```

Pivot Data Longer to have 1 response column

```{r}
nmdata_longer_dose_duration <- nmdata_longer_dose_duration %>% 
  tidyr::pivot_longer(cols = (UPR_ER_Stress: Cytoskeleton_Organization), 
                      names_to = "responses", 
               values_to = "effects")
```


Change Exposure dose to numeric class

```{r}
nmdata_longer_dose_duration$exposure_dose_mg_ml <- as.numeric(nmdata_longer_dose_duration$exposure_dose_mg_ml)
```

Plot Violin Plot

*** NOT a great way tp visualize the Data


```{r, warning=FALSE}
ggplot(data = nmdata_longer_dose_duration, aes(fill=effects, y=exposure_dose_mg_ml, x=responses)) + 
    geom_violin(position="dodge", alpha=0.5, outlier.colour="transparent") +
    scale_fill_viridis(discrete=T, name="") +
    theme_ipsum()  +
    xlab("Bioassay") +
    ylab("exposure dose mg/mL") +
    ggtitle("Influence of Nanoparticle Dose on Bioassay Outcome up to 30 mg/mL") +
  theme(plot.title = element_text(size = 10, hjust = 0.5)) +
    ylim(0,30) +
  coord_flip() 
```
Add Duration



```{r}
ggplot(data = nmdata_longer_dose_duration, aes(x = exposure_duration_h, y = exposure_dose_mg_ml, color = effects)) +
  geom_point()
```
```{r}

```



```{r}
ggplot(data = nmdata_longer_dose_duration, aes(x = exposure_duration_h,
                                               y = exposure_dose_mg_ml,
                                               color = nanoparticles)) +
  geom_point() +
  facet_wrap(~effects + responses) 
```


```{r}
nmdata_longer_dose_duration_metal_ion <- nmdata_longer_dose_duration %>% 
  filter(responses == "Metal_Ion")
```


```{r}
plot <- ggplot(data = nmdata_longer_dose_duration_metal_ion, aes(x = exposure_duration_h,
                                               y = exposure_dose_mg_ml,
                                               color = nanoparticles)) +
  geom_point() +
  facet_wrap(~effects + responses) 
```

```{r}
ggplotly(plot)
```


```{r}
nm_data_long_response <- nmdata %>% 
  pivot_longer(cols = Cell_Cycle_and_Proliferation:Cytoskeleton_Organization, names_to = "assay", values_to = "response")
```

```{r}
# interests <- nm_data_long_response %>%
#   group_by(nanoparticles, assay, exposure_dose_mg_ml, exposure_duration_h) %>%
#   filter(n_distinct(response) > 1) %>%
#   ungroup()
```

```{r}
# nm_data_long_response_low_exposure <- nm_data_long_response %>% 
#   filter(exposure_dose_mg_ml <= 1)
```

```{r}
nm_data_long_response$exposure_dose_mg_ml <- as.factor(nm_data_long_response$exposure_dose_mg_ml)
```



```{r}
nm_data_long_response$exposure_dose_mg_ml <- factor(nm_data_long_response$exposure_dose_mg_ml,
                                                   levels = c(0.2, 1, 2.5, 3, 5, 10, 12, 20, 25,
                                                              30, 40, 50, 60, 75, 100, 150, 200,
                                                              400, 1000))
```

```{r}
ggplot(data = nm_data_long_response, aes(x = exposure_dose_mg_ml,
                                         
                                         fill = response)) +
  geom_bar(stat = "count", position = "dodge", color="black") +
  facet_wrap(~assay)+
  theme_minimal()
  
```





