---
title: "random_forest_split"
author: "AJ Luna"
date: "2023-02-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#load libraries
if (!require(librarian)){
  install.packages("librarian")
  library(librarian)
}

librarian::shelf(here,
readr,
janitor,
tidyr,
tidyverse,
plotly,
randomForest,
edarf,
randomForestExplainer,
caret)

```

Load in Data


```{r}
#Import data
nmdata <- read_csv(here("data/Literature_transcriptomics_review_and_data_of_Nanoparticle_Induced_Cellular_Outcomes.csv"))

```

Clean Column Names


```{r}
nmdata <- nmdata %>% 
  clean_names()
```

Replace "?" with NA Values


```{r}
nmdata[nmdata == "?"] <- NA
```

Omit rows with NA Values (Dataset goes from 244 observations to 178)

```{r}
nmdata_no_na <- na.omit(nmdata)
```



Convert response column (column model is guessing) to class factor
```{r}
#cell cycle and profileration 
nmdata_no_na$cell_cycle_and_proliferation_responses <- as.factor(nmdata_no_na$cell_cycle_and_proliferation_responses)

#cell death and aoptosis
nmdata_no_na$cell_death_and_apoptosis_responses <- as.factor(nmdata_no_na$cell_death_and_apoptosis_responses)

#dna damage and repair
nmdata_no_na$dna_damage_and_repair_responses <- as.factor(nmdata_no_na$dna_damage_and_repair_responses)

#cell adhesion and extracelluar matrix response
nmdata_no_na$cell_adhesion_and_extracellular_matrix_responses <- as.factor(nmdata_no_na$cell_adhesion_and_extracellular_matrix_responses)

#inflammation and immune responses
nmdata_no_na$inflammation_and_immune_responses<- as.factor(nmdata_no_na$inflammation_and_immune_responses)

#protein responses and endoplastimic reticulum er stress
nmdata_no_na$unfolded_protein_responses_upr_and_endoplasmatic_reticulum_er_stress <- as.factor(nmdata_no_na$unfolded_protein_responses_upr_and_endoplasmatic_reticulum_er_stress)

#metal ion responses
nmdata_no_na$metal_ion_responses <- as.factor(nmdata_no_na$metal_ion_responses)

#angio genesis responses
nmdata_no_na$angiogenenesis_responses <- as.factor(nmdata_no_na$angiogenenesis_responses)

#cytoskeleton organization responses
nmdata_no_na$cytoskeleton_organization_responses<- as.factor(nmdata_no_na$cytoskeleton_organization_responses)


```






Randomize and Split Data 
```{r}
# Set the seed for reproducibility
set.seed(123)

# Create a random index vector to split the data
index <- sample(2, nrow(nmdata_no_na), replace = TRUE, prob = c(0.6, 0.4))

# Split the data into training and testing sets
train_data <- nmdata_no_na[index == 1, ]
test_data <- nmdata_no_na[index == 2, ]

```

Random Forest Cell Cycle Proliferation
```{r}

# Train the random forest model on the training set
rf_cell_cycle_and_proliferation_responses_split <- randomForest(cell_cycle_and_proliferation_responses ~ ., data = train_data, importance = TRUE,proximity = TRUE)

rf_cell_cycle_and_proliferation_responses_split

# Make predictions on the testing set using the trained model
predictions_cellcycle <- predict(rf_cell_cycle_and_proliferation_responses_split, test_data)

predictions_cellcycle

# Evaluate the performance of the model on the testing set
performance_cellcycle <- confusionMatrix(predictions_cellcycle,test_data$cell_cycle_and_proliferation_responses)

performance_cellcycle
```

Variable of Importance Cell Cycle Proliferation
```{r}
#Cell Cycle Proliferation Variables of Importance
ImpData_rf_cell_cycle_and_proliferation_responses_split <- as.data.frame(importance(rf_cell_cycle_and_proliferation_responses_split))
ImpData_rf_cell_cycle_and_proliferation_responses_split$Var.Names <- row.names(ImpData_rf_cell_cycle_and_proliferation_responses_split)

ImpData_rf_cell_cycle_and_proliferation_responses_split <- ImpData_rf_cell_cycle_and_proliferation_responses_split %>% mutate(Var.Names = fct_reorder(Var.Names, .[['MeanDecreaseAccuracy']]))

ggplot(ImpData_rf_cell_cycle_and_proliferation_responses_split, aes(x=Var.Names, y=`MeanDecreaseAccuracy`)) +
  geom_segment( aes(x=Var.Names, xend=Var.Names, y=0, yend=`MeanDecreaseAccuracy`), color="skyblue") +
  geom_point(aes(size = MeanDecreaseGini), color="blue", alpha=0.6) +
  theme_light() +
  coord_flip() +
  theme(
    legend.position="bottom",
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank()
  )

```


Random Forest Cell Death and Apoptosis 
```{r}

# Train the random forest model on the training set
rf_cell_death_and_apoptosis_responses_split <- randomForest(cell_death_and_apoptosis_responses ~ ., data = train_data, importance = TRUE,proximity = TRUE)

rf_cell_death_and_apoptosis_responses_split

# Make predictions on the testing set using the trained model
predictions_celldeath <- predict(rf_cell_death_and_apoptosis_responses_split, test_data)

predictions_celldeath

# Evaluate the performance of the model on the testing set
performance_celldeath <- confusionMatrix(predictions_celldeath,test_data$cell_death_and_apoptosis_responses)

performance_celldeath

```

Variables of Importance Cell Death and Apoptosis
```{r}

#Cell Death and Apoptosis Variables of Importance
ImpData_rf_cell_death_and_apoptosis_responses_split <- as.data.frame(importance(rf_cell_death_and_apoptosis_responses_split))
ImpData_rf_cell_death_and_apoptosis_responses_split$Var.Names <- row.names(ImpData_rf_cell_death_and_apoptosis_responses_split)

ImpData_rf_cell_death_and_apoptosis_responses_split <- ImpData_rf_cell_death_and_apoptosis_responses_split %>% mutate(Var.Names = fct_reorder(Var.Names, .[['MeanDecreaseAccuracy']]))

ggplot(ImpData_rf_cell_death_and_apoptosis_responses_split, aes(x=Var.Names, y=`MeanDecreaseAccuracy`)) +
  geom_segment( aes(x=Var.Names, xend=Var.Names, y=0, yend=`MeanDecreaseAccuracy`), color="skyblue") +
  geom_point(aes(size = MeanDecreaseGini), color="blue", alpha=0.6) +
  theme_light() +
  coord_flip() +
  theme(
    legend.position="bottom",
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank()
  )

```


Randomforest DNA Damage and Repair Response
```{r}
# Train the random forest model on the training set
rf_dna_damage_and_repair_responses_split <- randomForest(dna_damage_and_repair_responses ~ ., data = train_data, importance = TRUE,proximity = TRUE)

rf_dna_damage_and_repair_responses_split

# Make predictions on the testing set using the trained model
predictions_dna_damage <- predict(rf_dna_damage_and_repair_responses_split, test_data)

predictions_dna_damage

# Evaluate the performance of the model on the testing set
performance_dna_damage <- confusionMatrix(predictions_dna_damage,test_data$dna_damage_and_repair_responses)

performance_dna_damage

```


Variables of Importance DNA Damage and Repair
```{r}
#DNA Damage and Repair Variables of Importance
ImpData_rf_dna_damage_and_repair_responses_split <- as.data.frame(importance(rf_dna_damage_and_repair_responses_split))
ImpData_rf_dna_damage_and_repair_responses_split$Var.Names <- row.names(ImpData_rf_dna_damage_and_repair_responses_split)

ImpData_rf_dna_damage_and_repair_responses_split <- ImpData_rf_dna_damage_and_repair_responses_split %>% mutate(Var.Names = fct_reorder(Var.Names, .[['MeanDecreaseAccuracy']]))

ggplot(ImpData_rf_dna_damage_and_repair_responses_split, aes(x=Var.Names, y=`MeanDecreaseAccuracy`)) +
  geom_segment( aes(x=Var.Names, xend=Var.Names, y=0, yend=`MeanDecreaseAccuracy`), color="skyblue") +
  geom_point(aes(size = MeanDecreaseGini), color="blue", alpha=0.6) +
  theme_light() +
  coord_flip() +
  theme(
    legend.position="bottom",
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank()
  )

```


Random Forest Cell Adhesion and Extracelluar Matrix 
```{r}
# Train the random forest model on the training set
rf_cell_adhesion_and_extracellular_matrix_responses_split <- randomForest(cell_adhesion_and_extracellular_matrix_responses ~ ., data = train_data, importance = TRUE,proximity = TRUE)

rf_cell_adhesion_and_extracellular_matrix_responses_split

# Make predictions on the testing set using the trained model
predictions_cell_adhesion <- predict(rf_cell_adhesion_and_extracellular_matrix_responses_split, test_data)

predictions_cell_adhesion

# Evaluate the performance of the model on the testing set
performance_cell_adhesion <- confusionMatrix(predictions_cell_adhesion,test_data$cell_adhesion_and_extracellular_matrix_responses)

performance_cell_adhesion

```


Variable of Importance Cell Adhesion and Extracelluar Matrix
```{r}

#Cell Adhesion and Extracelluar  Variables of Importance
ImpData_rf_cell_adhesion_and_extracellular_matrix_responses_split <- as.data.frame(importance(rf_cell_adhesion_and_extracellular_matrix_responses_split))
ImpData_rf_cell_adhesion_and_extracellular_matrix_responses_split$Var.Names <- row.names(ImpData_rf_cell_adhesion_and_extracellular_matrix_responses_split)

ImpData_rf_cell_adhesion_and_extracellular_matrix_responses_split <- ImpData_rf_cell_adhesion_and_extracellular_matrix_responses_split %>% mutate(Var.Names = fct_reorder(Var.Names, .[['MeanDecreaseAccuracy']]))

ggplot(ImpData_rf_cell_adhesion_and_extracellular_matrix_responses_split, aes(x=Var.Names, y=`MeanDecreaseAccuracy`)) +
  geom_segment( aes(x=Var.Names, xend=Var.Names, y=0, yend=`MeanDecreaseAccuracy`), color="skyblue") +
  geom_point(aes(size = MeanDecreaseGini), color="blue", alpha=0.6) +
  theme_light() +
  coord_flip() +
  theme(
    legend.position="bottom",
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank()
  )


```


Random Forest Inflammation and Immune Responses
```{r}
# Train the random forest model on the training set
rf_inflammation_and_immune_responses_split <- randomForest(inflammation_and_immune_responses ~ ., data = train_data, importance = TRUE,proximity = TRUE)

rf_inflammation_and_immune_responses_split

# Make predictions on the testing set using the trained model
predictions_inflammation_and_immune <- predict(rf_inflammation_and_immune_responses_split, test_data)

predictions_inflammation_and_immune

# Evaluate the performance of the model on the testing set
performance_inflammation_and_immune <- confusionMatrix(predictions_inflammation_and_immune,test_data$inflammation_and_immune_responses)

performance_inflammation_and_immune

```

Variables of Importance Inflammation and Immune Response 
```{r}
#Inflammation and Immune Variables of Importance
ImpData_rf_inflammation_and_immune_responses_split <- as.data.frame(importance(rf_inflammation_and_immune_responses_split))
ImpData_rf_inflammation_and_immune_responses_split$Var.Names <- row.names(ImpData_rf_inflammation_and_immune_responses_split)

ImpData_rf_inflammation_and_immune_responses_split <- ImpData_rf_inflammation_and_immune_responses_split %>% mutate(Var.Names = fct_reorder(Var.Names, .[['MeanDecreaseAccuracy']]))

ggplot(ImpData_rf_inflammation_and_immune_responses_split, aes(x=Var.Names, y=`MeanDecreaseAccuracy`)) +
  geom_segment( aes(x=Var.Names, xend=Var.Names, y=0, yend=`MeanDecreaseAccuracy`), color="skyblue") +
  geom_point(aes(size = MeanDecreaseGini), color="blue", alpha=0.6) +
  theme_light() +
  coord_flip() +
  theme(
    legend.position="bottom",
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank()
  )

```


Random Forest Metal Ions
```{r}
# Train the random forest model on the training set
rf_metal_ion_responses_split <- randomForest(metal_ion_responses ~ ., data = train_data, importance = TRUE,proximity = TRUE)

rf_metal_ion_responses_split

# Make predictions on the testing set using the trained model
predictions_metal_ion <- predict(rf_metal_ion_responses_split, test_data)

predictions_metal_ion

# Evaluate the performance of the model on the testing set
performance_metal_ion <- confusionMatrix(predictions_metal_ion,test_data$metal_ion_responses)

performance_metal_ion

```




Variable of Importance Metal Ions
```{r}
#Metal Ions  Variables of Importance
ImpData_rf_metal_ion_responses_split <- as.data.frame(importance(rf_metal_ion_responses_split))
ImpData_rf_metal_ion_responses_split$Var.Names <- row.names(ImpData_rf_metal_ion_responses_split)

ImpData_rf_metal_ion_responses_split <- ImpData_rf_metal_ion_responses_split %>% mutate(Var.Names = fct_reorder(Var.Names, .[['MeanDecreaseAccuracy']]))

ggplot(ImpData_rf_metal_ion_responses_split, aes(x=Var.Names, y=`MeanDecreaseAccuracy`)) +
  geom_segment( aes(x=Var.Names, xend=Var.Names, y=0, yend=`MeanDecreaseAccuracy`), color="skyblue") +
  geom_point(aes(size = MeanDecreaseGini), color="blue", alpha=0.6) +
  theme_light() +
  coord_flip() +
  theme(
    legend.position="bottom",
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank()
  )

```

Random Forest Unfolded Protein 
```{r}

# Train the random forest model on the training set
rf_unfolded_protein_responses_upr_and_endoplasmatic_reticulum_er_stress_split <- randomForest(unfolded_protein_responses_upr_and_endoplasmatic_reticulum_er_stress ~ ., data = train_data, importance = TRUE,proximity = TRUE)

rf_unfolded_protein_responses_upr_and_endoplasmatic_reticulum_er_stress_split

# Make predictions on the testing set using the trained model
predictions_unfolded_protein <- predict(rf_unfolded_protein_responses_upr_and_endoplasmatic_reticulum_er_stress_split, test_data)

predictions_unfolded_protein

# Evaluate the performance of the model on the testing set
performance_unfolded_protein <- confusionMatrix(predictions_unfolded_protein,test_data$unfolded_protein_responses_upr_and_endoplasmatic_reticulum_er_stress)

performance_unfolded_protein

```



Variable of Importance Unfolded Protein and Endplasmatic Reticulum
```{r}
#Metal Ions  Variables of Importance
ImpData_rf_unfolded_protein_responses_upr_and_endoplasmatic_reticulum_er_stress_split <- as.data.frame(importance(rf_unfolded_protein_responses_upr_and_endoplasmatic_reticulum_er_stress_split))
ImpData_rf_unfolded_protein_responses_upr_and_endoplasmatic_reticulum_er_stress_split$Var.Names <- row.names(ImpData_rf_unfolded_protein_responses_upr_and_endoplasmatic_reticulum_er_stress_split)

ImpData_rf_unfolded_protein_responses_upr_and_endoplasmatic_reticulum_er_stress_split <- ImpData_rf_unfolded_protein_responses_upr_and_endoplasmatic_reticulum_er_stress_split %>% mutate(Var.Names = fct_reorder(Var.Names, .[['MeanDecreaseAccuracy']]))

ggplot(ImpData_rf_unfolded_protein_responses_upr_and_endoplasmatic_reticulum_er_stress_split, aes(x=Var.Names, y=`MeanDecreaseAccuracy`)) +
  geom_segment( aes(x=Var.Names, xend=Var.Names, y=0, yend=`MeanDecreaseAccuracy`), color="skyblue") +
  geom_point(aes(size = MeanDecreaseGini), color="blue", alpha=0.6) +
  theme_light() +
  coord_flip() +
  theme(
    legend.position="bottom",
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank()
  )


```





Facet Wrap Plot of variables of importance
```{r}
# Create "Assay" column to join data

ImpData_rf_metal_ion_responses_split$assay = "metal_ion_responses"
ImpData_rf_inflammation_and_immune_responses_split$assay = "inflammation_and_immune_responses"
ImpData_rf_cell_adhesion_and_extracellular_matrix_responses_split$assay = "cell_adhesion_and_extracellular_matrix_responses"
ImpData_rf_cell_death_and_apoptosis_responses_split$assay = "cell_death_and_apoptosis_responses"
ImpData_rf_cell_cycle_and_proliferation_responses_split$assay = "cell_cycle_and_proliferation_responses"
ImpData_rf_unfolded_protein_responses_upr_and_endoplasmatic_reticulum_er_stress_split$assay = "unfolded_protein_responses_upr_and_endoplasmatic_reticulum_er_stress"

```

Combine Data of All Variables of Importance
```{r}

#combine all data
df_list_good_models_split <- list(ImpData_rf_metal_ion_responses_split,
             ImpData_rf_inflammation_and_immune_responses_split, 
             ImpData_rf_cell_adhesion_and_extracellular_matrix_responses_split,
             ImpData_rf_cell_death_and_apoptosis_responses_split,
             ImpData_rf_cell_cycle_and_proliferation_responses_split,
             ImpData_rf_unfolded_protein_responses_upr_and_endoplasmatic_reticulum_er_stress_split)

```


```{r}
all_data_good_models_split <-  Reduce(function(x, y) merge(x, y, all=TRUE), df_list_good_models_split, accumulate=FALSE)

```


Plot All Variables of Importance
```{r}

plot2 <- ggplot(all_data_good_models_split, aes(x=Var.Names, y=`MeanDecreaseAccuracy`)) +
  geom_segment( aes(x=Var.Names, xend=Var.Names, y=0, yend=`MeanDecreaseAccuracy`), color="skyblue") +
  geom_point(aes(size = MeanDecreaseGini), color="blue", alpha=0.6) +
  theme_light() +
  coord_flip() +
  theme(
    legend.position="bottom",
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank()
  ) +
  facet_wrap(~assay)

ggplotly(plot2)

```


Random Forest Explainer Cell Cycle and proliferation
```{r}
# explain_forest(rf_cell_cycle_and_proliferation_responses_split, interactions = TRUE, data = nmdata_no_na, path = here("random_forests_explained/split_cell_cycle_and_proliferation_explained.html"))

```


Plot Prediction Interaction Cell Cycle Proliferation
```{r}
plot_predict_interaction(rf_cell_cycle_and_proliferation_responses_split, nmdata_no_na, "exposure_dose_mg_ml", "zeta_potential_m_v")

```


Random Forest Explainer Metal Ion
```{r}
# explain_forest(rf_metal_ion_responses_split, interactions = TRUE, data = nmdata_no_na, path = here("random_forests_explained/split_metal_ion_explained.html"))

```

Plot Prediction Interactions Metal Ion 
```{r}
plot_predict_interaction(rf_metal_ion_responses_split, nmdata_no_na, "exposure_dose_mg_ml", "zeta_potential_m_v")

```


Random Forest Explainer Cell Adhesion 
```{r}
# explain_forest(rf_cell_adhesion_and_extracellular_matrix_responses_split, interactions = TRUE, data = nmdata_no_na, path = here("random_forests_explained/split_cell_adhesion_and_extracellular_matrix_explained.html"))
```

Plot Prediction Interactions Cell adhesion
```{r}
plot_predict_interaction(rf_cell_adhesion_and_extracellular_matrix_responses_split, nmdata_no_na, "exposure_dose_mg_ml", "zeta_potential_m_v")

```


Random Forest Explainer Cell Death and Apoptosis
```{r}

# explain_forest(rf_cell_death_and_apoptosis_responses_split, interactions = TRUE, data = nmdata_no_na, path = here("random_forests_explained/split_cell_death_and_apoptosis_explained.html"))

```

Plot Prediction Interactions Cell Death and Apoptosis
```{r}

plot_predict_interaction(rf_cell_death_and_apoptosis_responses_split, nmdata_no_na, "exposure_dose_mg_ml", "zeta_potential_m_v")

```

Random Forest Explainer Inflammation and Immune 
```{r}
# explain_forest(rf_inflammation_and_immune_responses_split, interactions = TRUE, data = nmdata_no_na, path = here("random_forests_explained/split_inflammation_and_immune_explained.html"))
```


Plot Prediction Interactions Imflammation and Immune

```{r}
plot_predict_interaction(rf_inflammation_and_immune_responses_split, nmdata_no_na, "exposure_dose_mg_ml", "zeta_potential_m_v")

```


Random Forest Explainer Unfolded Protein and Endoplasmatic Reticulum
```{r}
# explain_forest(rf_unfolded_protein_responses_upr_and_endoplasmatic_reticulum_er_stress_split, interactions = TRUE, data = nmdata_no_na, path = here("random_forests_explained/split_unfolded_protein_responses_upr_and_endoplasmatic_reticulum_er_stress_explained.html"))

```


Plot Prediction Interactions Inflammation and Immune
```{r}

plot_predict_interaction(rf_unfolded_protein_responses_upr_and_endoplasmatic_reticulum_er_stress_split, nmdata_no_na, "exposure_dose_mg_ml", "zeta_potential_m_v")

```


