---
title: "Random Forest & Bayesian Network Comparison"
author: "Armando Luna, Brittany Johnson, Connor Flynn"
date: "2/3/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#load libraries
if (!require(librarian)){
  install.packages("librarian")
  library(librarian)
}

librarian::shelf(here,
readr,
janitor,
tidyr,
tidyverse,
plotly,
randomForest,
edarf,
randomForestExplainer)

```

The purpose of this R Markdown is to utilize a random forest classification algorithm to demonstrate the most important factors in predicting whether a bioassay will be triggered or not


Load in Data


```{r}
#Import data
nmdata <- read_csv(here("data/Literature_transcriptomics_review_and_data_of_Nanoparticle_Induced_Cellular_Outcomes.csv"))

```

Clean Column Names


```{r}
nmdata <- nmdata %>% 
  clean_names()
```

Replace "?" with NA Values


```{r}
nmdata[nmdata == "?"] <- NA
```

Drop microarray method transriptomics

```{r}
nmdata <- nmdata %>% 
  select(-microarray_method_transcriptomics)
```





Identify numeric data
```{r}
nmdata$exposure_dose_mg_ml <- as.numeric(nmdata$exposure_dose_mg_ml)
nmdata$exposure_duration_h <- as.numeric(nmdata$exposure_duration_h)
```



Omit rows with NA Values (Dataset goes from 244 observations to 178)

```{r}
nmdata_no_na <- na.omit(nmdata)
```



Convert response column (column model is guessing) to class factor
```{r}
#cell cycle and profileration 
nmdata_no_na$cell_cycle_and_proliferation_responses <- as.factor(nmdata_no_na$cell_cycle_and_proliferation_responses)

#cell death and aoptosis
nmdata_no_na$cell_death_and_apoptosis_responses <- as.factor(nmdata_no_na$cell_death_and_apoptosis_responses)

#dna damage and repair
nmdata_no_na$dna_damage_and_repair_responses <- as.factor(nmdata_no_na$dna_damage_and_repair_responses)

#cell adhesion and extracelluar matrix response
nmdata_no_na$cell_adhesion_and_extracellular_matrix_responses <- as.factor(nmdata_no_na$cell_adhesion_and_extracellular_matrix_responses)

#inflammation and immune responses
nmdata_no_na$inflammation_and_immune_responses<- as.factor(nmdata_no_na$inflammation_and_immune_responses)

#protein responses and endoplastimic reticulum er stress
nmdata_no_na$unfolded_protein_responses_upr_and_endoplasmatic_reticulum_er_stress <- as.factor(nmdata_no_na$unfolded_protein_responses_upr_and_endoplasmatic_reticulum_er_stress)

#metal ion responses
nmdata_no_na$metal_ion_responses <- as.factor(nmdata_no_na$metal_ion_responses)

#angio genesis responses
nmdata_no_na$angiogenenesis_responses <- as.factor(nmdata_no_na$angiogenenesis_responses)

#cytoskeleton organization responses
nmdata_no_na$cytoskeleton_organization_responses<- as.factor(nmdata_no_na$cytoskeleton_organization_responses)


```

Subset Data

```{r}
cell_cycle_and_proliferation_responses_df <- nmdata_no_na %>% 
  select(1:12)

cell_death_and_apoptosis_responses_df <- nmdata_no_na %>% 
  select(1:11, 13)

dna_damage_and_repair_responses_df <- nmdata_no_na %>% 
  select(1:11, 14)

cell_adhesion_and_extracellular_matrix_responses_df <- nmdata_no_na %>% 
  select(1:11, 15)

inflammation_and_immune_responses_df <-  nmdata_no_na %>% 
  select(1:11, 16)

unfolded_protein_responses_upr_and_endoplasmatic_reticulum_er_stress_df <-
  nmdata_no_na %>% 
  select(1:11, 17)

metal_ion_responses_df <- nmdata_no_na %>% 
  select(1:11, 18)

angiogenenesis_responses_df <- nmdata_no_na %>% 
  select(1:11, 19)

cytoskeleton_organization_responses_df <- nmdata_no_na %>% 
  select(1:11, 20)

```


Create random forest model to predict triggered vs no effect 
```{r}
#Create random forest model to predict triggered vs no effect in cell_cycle_and_proliferation_responses

rf_cell_cycle_and_proliferation_responses_limited <- randomForest(cell_cycle_and_proliferation_responses ~ ., 
                       data = cell_cycle_and_proliferation_responses_df, 
                       importance = TRUE,
                       proximity = TRUE)

rf_cell_cycle_and_proliferation_responses_limited 

```


```{r}
#Create random forest model to predict triggered vs no effect in cell_death_and_apoptosis_responses

rf_cell_death_and_apoptosis_responses_limited <- randomForest(cell_death_and_apoptosis_responses ~ ., 
                       data = cell_death_and_apoptosis_responses_df, 
                       importance = TRUE,
                       proximity = TRUE)

rf_cell_death_and_apoptosis_responses_limited

```


```{r}
#Create random forest model to predict triggered vs no effect in dna_damage_and_repair_responses

rf_dna_damage_and_repair_responses_limited <- randomForest(dna_damage_and_repair_responses ~ ., 
                       data = dna_damage_and_repair_responses_df, 
                       importance = TRUE,
                       proximity = TRUE)

rf_dna_damage_and_repair_responses_limited

```


```{r}

#Create random forest model to predict triggered vs no effect in cell_adhesion_and_extracellular_matrix_responses

rf_cell_adhesion_and_extracellular_matrix_responses_limited <- randomForest(cell_adhesion_and_extracellular_matrix_responses ~ ., 
                       data = cell_adhesion_and_extracellular_matrix_responses_df, 
                       importance = TRUE,
                       proximity = TRUE)

rf_cell_adhesion_and_extracellular_matrix_responses_limited

```


```{r}
#Create random forest model to predict triggered vs no effect in inflammation_and_immune_responses

rf_inflammation_and_immune_responses_limited <- randomForest(inflammation_and_immune_responses ~ ., 
                       data = inflammation_and_immune_responses_df, 
                       importance = TRUE,
                       proximity = TRUE)

rf_inflammation_and_immune_responses_limited

```


```{r}
#Create random forest model to predict triggered vs no effect in unfolded_protein_responses

rf_unfolded_protein_responses_upr_and_endoplasmatic_reticulum_er_stress__responses_limited <- randomForest(unfolded_protein_responses_upr_and_endoplasmatic_reticulum_er_stress ~ ., 
                       data = unfolded_protein_responses_upr_and_endoplasmatic_reticulum_er_stress_df, 
                       importance = TRUE,
                       proximity = TRUE)

rf_unfolded_protein_responses_upr_and_endoplasmatic_reticulum_er_stress__responses_limited


```


```{r}

#Create random forest model to predict triggered vs no effect in metal_ion_responses

rf_metal_ion_responses_limited <- randomForest(metal_ion_responses ~ ., 
                       data = metal_ion_responses_df, 
                       importance = TRUE,
                       proximity = TRUE)

rf_metal_ion_responses_limited

```



```{r}

#Create random forest model to predict triggered vs no effect in angiogenenesis_responses

rf_angiogenenesis_responses_limited <- randomForest(angiogenenesis_responses ~ ., 
                       data = angiogenenesis_responses_df, 
                       importance = TRUE,
                       proximity = TRUE)

rf_angiogenenesis_responses_limited
```


```{r}

#Create random forest model to predict triggered vs no effect in cytoskeleton_organization_responses

rf_cytoskeleton_organization_responses_limited <- randomForest(cytoskeleton_organization_responses ~ ., 
                       data = cytoskeleton_organization_responses_df, 
                       importance = TRUE,
                       proximity = TRUE)

rf_cytoskeleton_organization_responses_limited

```


Variables of Importance
```{r}
#cell cycle and proliferation variable of importances
ImpData_rf_cell_cycle_and_proliferation_responses_limited <- as.data.frame(importance(rf_cell_cycle_and_proliferation_responses_limited))
ImpData_rf_cell_cycle_and_proliferation_responses_limited$Var.Names <- row.names(ImpData_rf_cell_cycle_and_proliferation_responses_limited)

ImpData_rf_cell_cycle_and_proliferation_responses_limited <- ImpData_rf_cell_cycle_and_proliferation_responses_limited %>% mutate(Var.Names = fct_reorder(Var.Names, .[['MeanDecreaseAccuracy']]))

ggplot(ImpData_rf_cell_cycle_and_proliferation_responses_limited, aes(x=Var.Names, y=`MeanDecreaseAccuracy`)) +
  geom_segment( aes(x=Var.Names, xend=Var.Names, y=0, yend=`MeanDecreaseAccuracy`), color="skyblue") +
  geom_point(aes(size = MeanDecreaseGini), color="blue", alpha=0.6) +
  theme_light() +
  coord_flip() +
  theme(
    legend.position="bottom",
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank()
  )

```


```{r}

#cell_death_and_apoptosis_responses_limited variable of importances
ImpData_rf_cell_death_and_apoptosis_responses_limited <- as.data.frame(importance(rf_cell_death_and_apoptosis_responses_limited))
ImpData_rf_cell_death_and_apoptosis_responses_limited$Var.Names <- row.names(ImpData_rf_cell_death_and_apoptosis_responses_limited)

ImpData_rf_cell_death_and_apoptosis_responses_limited <- ImpData_rf_cell_death_and_apoptosis_responses_limited %>% mutate(Var.Names = fct_reorder(Var.Names, .[['MeanDecreaseAccuracy']]))

ggplot(ImpData_rf_cell_death_and_apoptosis_responses_limited, aes(x=Var.Names, y=`MeanDecreaseAccuracy`)) +
  geom_segment( aes(x=Var.Names, xend=Var.Names, y=0, yend=`MeanDecreaseAccuracy`), color="skyblue") +
  geom_point(aes(size = MeanDecreaseGini), color="blue", alpha=0.6) +
  theme_light() +
  coord_flip() +
  theme(
    legend.position="bottom",
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank()
  )

```


```{r}

#dna_damage_and_repair_responses_limited variable of importances
ImpData_rf_dna_damage_and_repair_responses_limited <- as.data.frame(importance(rf_dna_damage_and_repair_responses_limited))
ImpData_rf_dna_damage_and_repair_responses_limited$Var.Names <- row.names(ImpData_rf_dna_damage_and_repair_responses_limited)

ImpData_rf_dna_damage_and_repair_responses_limited <- ImpData_rf_dna_damage_and_repair_responses_limited %>% mutate(Var.Names = fct_reorder(Var.Names, .[['MeanDecreaseAccuracy']]))

ggplot(ImpData_rf_dna_damage_and_repair_responses_limited, aes(x=Var.Names, y=`MeanDecreaseAccuracy`)) +
  geom_segment( aes(x=Var.Names, xend=Var.Names, y=0, yend=`MeanDecreaseAccuracy`), color="skyblue") +
  geom_point(aes(size = MeanDecreaseGini), color="blue", alpha=0.6) +
  theme_light() +
  coord_flip() +
  theme(
    legend.position="bottom",
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank()
  )

```


```{r}

#cell_adhesion_and_extracellular_matrix_responses_limited variable of importances
ImpData_rf_cell_adhesion_and_extracellular_matrix_responses_limited <- as.data.frame(importance(rf_cell_adhesion_and_extracellular_matrix_responses_limited))
ImpData_rf_cell_adhesion_and_extracellular_matrix_responses_limited$Var.Names <- row.names(ImpData_rf_cell_adhesion_and_extracellular_matrix_responses_limited)

ImpData_rf_cell_adhesion_and_extracellular_matrix_responses_limited <- ImpData_rf_cell_adhesion_and_extracellular_matrix_responses_limited %>% mutate(Var.Names = fct_reorder(Var.Names, .[['MeanDecreaseAccuracy']]))

ggplot(ImpData_rf_cell_adhesion_and_extracellular_matrix_responses_limited, aes(x=Var.Names, y=`MeanDecreaseAccuracy`)) +
  geom_segment( aes(x=Var.Names, xend=Var.Names, y=0, yend=`MeanDecreaseAccuracy`), color="skyblue") +
  geom_point(aes(size = MeanDecreaseGini), color="blue", alpha=0.6) +
  theme_light() +
  coord_flip() +
  theme(
    legend.position="bottom",
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank()
  )

```


```{r}

#inflammation_and_immune_responses_limited variable of importances
ImpData_rf_inflammation_and_immune_responses_limited <- as.data.frame(importance(rf_inflammation_and_immune_responses_limited))
ImpData_rf_inflammation_and_immune_responses_limited$Var.Names <- row.names(ImpData_rf_inflammation_and_immune_responses_limited)

ImpData_rf_inflammation_and_immune_responses_limited <- ImpData_rf_inflammation_and_immune_responses_limited %>% mutate(Var.Names = fct_reorder(Var.Names, .[['MeanDecreaseAccuracy']]))

ggplot(ImpData_rf_inflammation_and_immune_responses_limited, aes(x=Var.Names, y=`MeanDecreaseAccuracy`)) +
  geom_segment( aes(x=Var.Names, xend=Var.Names, y=0, yend=`MeanDecreaseAccuracy`), color="skyblue") +
  geom_point(aes(size = MeanDecreaseGini), color="blue", alpha=0.6) +
  theme_light() +
  coord_flip() +
  theme(
    legend.position="bottom",
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank()
  )

```


```{r}

#unfolded_protein_responses_upr_and_endoplasmatic_reticulum_er_stress__responses_limited variable of importances
ImpData_rf_unfolded_protein_responses_upr_and_endoplasmatic_reticulum_er_stress__responses_limited <- as.data.frame(importance(rf_unfolded_protein_responses_upr_and_endoplasmatic_reticulum_er_stress__responses_limited))
ImpData_rf_unfolded_protein_responses_upr_and_endoplasmatic_reticulum_er_stress__responses_limited$Var.Names <- row.names(ImpData_rf_unfolded_protein_responses_upr_and_endoplasmatic_reticulum_er_stress__responses_limited)

ImpData_rf_unfolded_protein_responses_upr_and_endoplasmatic_reticulum_er_stress__responses_limited <- ImpData_rf_unfolded_protein_responses_upr_and_endoplasmatic_reticulum_er_stress__responses_limited %>% mutate(Var.Names = fct_reorder(Var.Names, .[['MeanDecreaseAccuracy']]))

ggplot(ImpData_rf_unfolded_protein_responses_upr_and_endoplasmatic_reticulum_er_stress__responses_limited, aes(x=Var.Names, y=`MeanDecreaseAccuracy`)) +
  geom_segment( aes(x=Var.Names, xend=Var.Names, y=0, yend=`MeanDecreaseAccuracy`), color="skyblue") +
  geom_point(aes(size = MeanDecreaseGini), color="blue", alpha=0.6) +
  theme_light() +
  coord_flip() +
  theme(
    legend.position="bottom",
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank()
  )

```



```{r}

#metal_ion_responses_limited variable of importances
ImpData_rf_metal_ion_responses_limited <- as.data.frame(importance(rf_metal_ion_responses_limited))
ImpData_rf_metal_ion_responses_limited$Var.Names <- row.names(ImpData_rf_metal_ion_responses_limited)

ImpData_rf_metal_ion_responses_limited <- ImpData_rf_metal_ion_responses_limited %>% mutate(Var.Names = fct_reorder(Var.Names, .[['MeanDecreaseAccuracy']]))

ggplot(ImpData_rf_metal_ion_responses_limited, aes(x=Var.Names, y=`MeanDecreaseAccuracy`)) +
  geom_segment( aes(x=Var.Names, xend=Var.Names, y=0, yend=`MeanDecreaseAccuracy`), color="skyblue") +
  geom_point(aes(size = MeanDecreaseGini), color="blue", alpha=0.6) +
  theme_light() +
  coord_flip() +
  theme(
    legend.position="bottom",
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank()
  )

```

```{r}

#angiogenenesis_responses_limited variable of importances
ImpData_rf_angiogenenesis_responses_limited <- as.data.frame(importance(rf_angiogenenesis_responses_limited))
ImpData_rf_angiogenenesis_responses_limited$Var.Names <- row.names(ImpData_rf_angiogenenesis_responses_limited)

ImpData_rf_angiogenenesis_responses_limited <- ImpData_rf_angiogenenesis_responses_limited %>% mutate(Var.Names = fct_reorder(Var.Names, .[['MeanDecreaseAccuracy']]))

ggplot(ImpData_rf_angiogenenesis_responses_limited, aes(x=Var.Names, y=`MeanDecreaseAccuracy`)) +
  geom_segment( aes(x=Var.Names, xend=Var.Names, y=0, yend=`MeanDecreaseAccuracy`), color="skyblue") +
  geom_point(aes(size = MeanDecreaseGini), color="blue", alpha=0.6) +
  theme_light() +
  coord_flip() +
  theme(
    legend.position="bottom",
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank()
  )

```


```{r}

#cytoskeleton_organization_responses_limited variable of importances
ImpData_rf_cytoskeleton_organization_responses_limited <- as.data.frame(importance(rf_cytoskeleton_organization_responses_limited
))
ImpData_rf_cytoskeleton_organization_responses_limited$Var.Names <- row.names(ImpData_rf_cytoskeleton_organization_responses_limited)

ImpData_rf_cytoskeleton_organization_responses_limited<- ImpData_rf_cytoskeleton_organization_responses_limited %>% mutate(Var.Names = fct_reorder(Var.Names, .[['MeanDecreaseAccuracy']]))

ggplot(ImpData_rf_cytoskeleton_organization_responses_limited, aes(x=Var.Names, y=`MeanDecreaseAccuracy`)) +
  geom_segment( aes(x=Var.Names, xend=Var.Names, y=0, yend=`MeanDecreaseAccuracy`), color="skyblue") +
  geom_point(aes(size = MeanDecreaseGini), color="blue", alpha=0.6) +
  theme_light() +
  coord_flip() +
  theme(
    legend.position="bottom",
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank()
  )

```



Add column to join data
```{r}

#add assay column to be able to join data
ImpData_rf_metal_ion_responses_limited$assay = "metal_ion_responses"
ImpData_rf_inflammation_and_immune_responses_limited$assay = "inflammation_and_immune_responses"
#ImpData_rf_dna_damage_and_repair_responses_limited$assay = "dna_damage_and_repair_responses"
#ImpData_rf_cytoskeleton_organization_responses_limited$assay = "cytoskeleton_organization_responses"
ImpData_rf_cell_adhesion_and_extracellular_matrix_responses_limited$assay = "cell_adhesion_and_extracellular_matrix_responses"
ImpData_rf_cell_death_and_apoptosis_responses_limited$assay = "cell_death_and_apoptosis_responses"
ImpData_rf_cell_cycle_and_proliferation_responses_limited$assay = "cell_cycle_and_proliferation_responses"
ImpData_rf_unfolded_protein_responses_upr_and_endoplasmatic_reticulum_er_stress__responses_limited$assay = "unfolded_protein_responses_upr_and_endoplasmatic_reticulum_er_stress"

```


Facet Wrap Plots
```{r}

#combine all data
df_list_limited <- list(ImpData_rf_metal_ion_responses_limited,
             ImpData_rf_inflammation_and_immune_responses_limited, 
             #ImpData_rf_dna_damage_and_repair_responses_limited,
            # ImpData_rf_cytoskeleton_organization_responses_limited,
             ImpData_rf_cell_adhesion_and_extracellular_matrix_responses_limited,
             ImpData_rf_cell_death_and_apoptosis_responses_limited,
             ImpData_rf_cell_cycle_and_proliferation_responses_limited,
             ImpData_rf_unfolded_protein_responses_upr_and_endoplasmatic_reticulum_er_stress__responses_limited)

```


```{r}

all_data_limited <- all_data <- Reduce(function(x, y) merge(x, y, all=TRUE), df_list_limited, accumulate=FALSE)

```


```{r}
plot <- ggplot(all_data_limited, aes(x=Var.Names, y=`MeanDecreaseAccuracy`)) +
  geom_segment( aes(x=Var.Names, xend=Var.Names, y=0, yend=`MeanDecreaseAccuracy`), color="skyblue") +
  geom_point(aes(size = MeanDecreaseGini), color="blue", alpha=0.6) +
  theme_light() +
  coord_flip() +
  theme(
    legend.position="bottom",
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank()
  ) +
  facet_wrap(~assay)

plot


```




```{r}
plot_predict_interaction(rf_metal_ion_responses_limited, nmdata_no_na, "exposure_dose_mg_ml", "zeta_potential_m_v")

```


```{r}
exposure <- as.data.frame(table(nmdata_no_na$exposure_dose_mg_ml))
exposure
```

```{r}

```

Random Forest Explainer

Cell Cycle and proliferation

```{r}
#explain_forest(rf_cell_cycle_and_proliferation_responses_limited, interactions = TRUE, data = nmdata_no_na, path = here("random_forests_limited_explained/cell_cycle_and_proliferation_explained.html"))
```



```{r}
plot_predict_interaction(rf_cell_cycle_and_proliferation_responses_limited, nmdata_no_na, "exposure_dose_mg_ml", "zeta_potential_m_v")
```




Can't seem to find another continuous variable that is important with zeta potential regarding prediction interaction

Metal Ion Responses

```{r}
#explain_forest(rf_metal_ion_responses_limited, interactions = TRUE, data = nmdata_no_na, path = here("random_forests_limited_explained/metal_ion_explained.html"))
```


```{r}
plot_predict_interaction(rf_cell_cycle_and_proliferation_responses_limited, nmdata_no_na, "exposure_dose_mg_ml", "zeta_potential_m_v")
```

Cell Adhesion and Extra cellular Matrix Responses

```{r}
#explain_forest(rf_cell_adhesion_and_extracellular_matrix_responses_limited, interactions = TRUE, data = nmdata_no_na, path = here("random_forests_limited_explained/cell_adhesion_and_extracellular_matrix_explained.html"))
```

```{r}
plot_predict_interaction(rf_cell_adhesion_and_extracellular_matrix_responses_limited, nmdata_no_na, "zeta_potential_m_v", "exposure_duration_h")
```

Cell Death and Apoptosis Responses

```{r}
#explain_forest(rf_cell_death_and_apoptosis_responses_limited, interactions = TRUE, data = nmdata_no_na, path = here("random_forests_limited_explained/cell_death_and_apoptosis_explained.html"))
```

```{r}
plot_predict_interaction(rf_cell_death_and_apoptosis_responses_limited, nmdata_no_na, "exposure_dose_mg_ml", "zeta_potential_m_v")
```

Inflammation and Immune Responses

```{r}
#explain_forest(rf_inflammation_and_immune_responses_limited, interactions = TRUE, data = nmdata_no_na, path = here("random_forests_limited_explained/inflammation_and_immune_explained.html"))
```

```{r}
plot_predict_interaction(rf_inflammation_and_immune_responses_limited, nmdata_no_na, "exposure_dose_mg_ml", "zeta_potential_m_v")
```



Unfolded Protein and Endoplasmatic Reticulum Stress

```{r}
#explain_forest(rf_unfolded_protein_responses_upr_and_endoplasmatic_reticulum_er_stress__responses_limited, interactions = TRUE, data = nmdata_no_na, path = here("random_forests_limited_explained/unfolded_protein_responses_upr_and_endoplasmatic_reticulum_er_stress_explained.html"))
```

